### 2018-10-05 figures/01_iris.Rmd
fill-out skeleton

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 6)
```

```{r read_chunks, cache=FALSE, echo=FALSE}
knitr::read_chunk("../scripts/99_utils.R")
```

```{r source_utils, message=FALSE, results='hide', echo=FALSE, warning=FALSE}

```

```{r load_data, message=FALSE, results='hide', echo=FALSE, warning=FALSE}
# setwd("figures")
source("../scripts/01.5_loaddata.R")
```

```{r 01_iris, echo=FALSE, warning=FALSE, fig.width = 5.5, fig.height = 3}
ggplot() + 
  geom_point(data = iris, aes(x = Sepal.Length, y = Sepal.Width))
```

### 2018-10-05 figures/A1_iris.Rmd
fill-out skeleton

--- 2018-10-05 figures/01_iris.Rmd
@@
 
-```{r 01_iris, echo=FALSE, warning=FALSE, fig.width = 5.5, fig.height = 3}
+```{r A1_iris, echo=FALSE, warning=FALSE, fig.width = 5.5, fig.height = 3}
 ggplot() + 
-  geom_point(data = iris, aes(x = Sepal.Length, y = Sepal.Width))
+  geom_point(data = iris, aes(x = Sepal.Length, y = Petal.Length))
 ```

### 2018-10-05 manuscript/appendix.Rmd
fill-out skeleton

---
title: "Appendix"
output:
  pdf_document:
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path = "../figures/", fig.width = 5, fig.height = 6)
```

```{r read_chunks, cache=FALSE, echo=FALSE}
knitr::read_chunk("../scripts/99_utils.R")
```

```{r source_utils, message=FALSE, results='hide', echo=FALSE, warning=FALSE}

```

```{r load_data, message=FALSE, results='hide', echo=FALSE, warning=FALSE}
  if(file.exists("../scripts/01.5_loaddata.R")){
    source("../scripts/01.5_loaddata.R")  
  }else{
    source("scripts/01.5_loaddata.R")  
  }
```

```{r A1_iris, fig.cap="Sepal length versus petal length", echo=FALSE, fig.width=1,out.height="4in"}
knitr::include_graphics("../figures/A1_iris-1.pdf")

```

### 2018-10-05 manuscript/figures.Rmd
fill-out skeleton

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 6, dev = c("pdf"))
```

```{r 01_iris, fig.cap="Sepal length versus width", echo=FALSE, fig.width=1,out.height="4in"}
knitr::include_graphics("../figures/01_iris-1.pdf")
```

### 2018-10-05 tables/01_iris_table.Rmd
fill-out skeleton

--- 2018-10-05 figures/01_iris.Rmd
@@
 ---
-title: "Figures"
 output:
-  pdf_document:
-    fig_caption: yes
+  pdf_document
+header-includes: 
+  - \pagenumbering{gobble}
 ---
@@
 ```{r load_data, message=FALSE, results='hide', echo=FALSE, warning=FALSE}
-# setwd("figures")
 source("../scripts/01.5_loaddata.R")
@@
 
-```{r 01_iris, echo=FALSE, warning=FALSE, fig.width = 5.5, fig.height = 3}
-ggplot() + 
-  geom_point(data = iris, aes(x = Sepal.Length, y = Sepal.Width))
+```{r iris_table, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
+knitr::kable(head(iris), caption = "Iris table")
 ```

### 2018-10-05 scripts/00_get_iris.R

@@ -1,0 +1,6 @@
+
+iris_path <- "data/iris.csv"
+# if(!file.exists(iris_path)){
+data("iris")
+write.csv(iris, iris_path, row.names = FALSE)
+# }

### 2018-10-05 scripts/01.5_loaddata.R

@@ -1,0 +1,1 @@
+iris <- read.csv("../data/iris.csv", stringsAsFactors = FALSE)

### 2018-10-05 scripts/01_prepdata.R

@@ -1,0 +1,7 @@
+# place any data preparation steps here
+
+library(tidyr)
+
+iris <- read.csv("data/iris.csv", stringsAsFactors = FALSE)
+iris_tidy_path <- "data/iris_tidy.csv"
+iris_tidy      <- tidyr::gather(iris, "variable", "value", -Species)

### 2018-10-05 scripts/02_modelling.R

@@ -1,0 +1,10 @@
+# setwd("scripts")
+source("99_utils.R")
+source("01_prepdata.R")
+
+# ---- load_packages ----
+library(broom)
+
+# ---- iris_lm ----
+fit <- lm(Sepal.Length ~ Petal.Length, data = iris)
+broom::tidy(fit)

### 2018-10-05 scripts/03_viz.R

@@ -1,0 +1,5 @@
+# setwd("scripts")
+# source("99_utils.R")
+# source("01_prepdata.R")
+
+# optionally construct plots here instead of in Rmd files

### 2018-10-05 scripts/04_tables.R

@@ -1,0 +1,5 @@
+# setwd("scripts")
+source("99_utils.R")
+# source("01_prepdata.R")
+
+# optionally construct tables here instead of in Rmd files

### 2018-10-05 scripts/99_utils.R

@@ -1,0 +1,23 @@
+
+# ---- source_utils ----
+
+library(ggplot2)
+
+theme_pred <- function(){
+  theme(legend.position = "na",
+        axis.text = element_text(size = 10),
+        axis.title = element_text(size = 12),
+        plot.title = element_text(size = 10, face = "bold", color = "black", hjust = 0))
+}
+
+signif_star <- function(x){
+  if(!is.na(x)){
+    if(x){
+      "*"
+    }else{
+      ""
+    }
+  }else{
+    ""
+  }
+}

### 2018-10-05 scripts/simulation.R
move simulation script from coursework

@@ -1,0 +1,195 @@
+# ---- pareto_demo ----
+library(dplyr)
+library(magrittr)
+library(ggplot2)
+library(cowplot)
+
+dpareto <- function(x, a=0.9, b=1) a*b^a/x^(a+1)
+ppareto <- function(x, a=0.9, b=1) (x > b)*(1-(b/x)^a)
+qpareto <- function(u, a=0.9, b=1) b/(1-u)^(1/a)
+rpareto <- function(n, a=0.9, b=1, max = Inf){
+  res <- qpareto(runif(n), a, b)
+  res <- res[res < max]
+  
+  while(length(res) < n){
+    res <- c(res, qpareto(runif(n - length(res)), a, b))
+  }
+  res
+} 
+# (dt <- sample(1:10, 12, replace = TRUE) %>% .[order(.)])
+# (cf <- cumulative_freq(dt))
+cumulative_freq <- function(x){
+  l_n    <- cumsum(rev(table(x)))
+  l_area <- rev(as.numeric(names(table(x))))
+  
+  data.frame(area = l_area, number = l_n)
+}
+
+# inv_cumulative_freq(cf)
+inv_cumulative_freq <- function(cf){
+  cf <- cf[order(cf$number),]
+  rep(cf$area, c(cf$number[1], diff(cf$number)))
+}
+
+# simulate random pareto draws
+# y <- rpareto(10000, max = 81935.7) # cap at the area of Lake Superior
+# saveRDS(y, "project/paper/data/y.rds")
+y <- readRDS("data/y.rds")
+
+plot_grid(
+  # individual binning
+  hist(log(y), plot = FALSE, n = 100) %>% 
+  {data.frame(x = .$breaks[-1], samples = log(.$counts))} %>%
+    ggplot + geom_line(aes(x, samples)) +
+    theme_minimal(),
+  # log binning
+  # hist(y, plot = FALSE, breaks = 10^(seq(from = 0, to = 6, by = 0.6))) %>% 
+  # {data.frame(x = .$breaks[-1], samples = .$counts + 0.01)} %>%
+  #   ggplot + geom_line(aes(x, samples)) +
+  #   theme_minimal() + scale_x_log10() + scale_y_log10(),
+  # cumulative binning
+  cumulative_freq(y) %>% 
+    ggplot + geom_line(aes(area, number)) +
+    theme_minimal() + scale_x_log10() + scale_y_log10() + 
[…]

### 2019-03-05 manuscript/manuscript.Rmd
fix file paths and makefile targets

---
title: Estimating Uncertainty in Global Lake Area

author:
  - name: Joseph Stachelek
    affiliation: a
address:
  - code: a
    address: Department of Fisheries and Wildlife, Michigan State University, East Lansing, MI, USA

lead_author_surname: Stachelek

# Place DOI URL or CRAN Package URL here
# doi: "https://cran.r-project.org/package=YourPackage"

abstract: | 
  Size is a critical factor determining the rate and occurrence of specific lake processes such as carbon sequestration. Unfortunately, the exact lake size-abundance distribution is unknown in part because we do not have a complete census of all lakes. More specifically, as lakes become smaller they are less likely to be included in lake databases either because they are too small to be resolved from remote sensing products or because of limited ground surveying effort. The exclusion of small lakes relative to large lakes is known as censoring.

  The present study explores one potential shortcoming of previous approaches at estimating global lake area and dealing with censoring. Namely, that the typical frequentist curve fitting and ad-hoc cutoff determination strategy (visual inspection to determine a likely censoring point) taken by previous studies yields an over-exact lake area estimate with no reported uncertainty. I address this shortcoming by fitting models in a Bayesian framework where each parameter contributes uncertainty to model estimates. I show that such although such models produce a more realistic estimate of lake area uncertainty they underestimate true total lake area. The degree of this underestimation is likely related to the proportion of the dataset subject to censoring. Ultimately, this may explain the fact that total lake area estimates have increased through time as the resolution of lake databases has improved.

# Optional: Acknowledgements
# acknowledgements: |

# Optional: One or more keywords
# keywords:
#   - one
#   - two
#   - optional
#   - keywords
#   - here

# Font size of the document, values of 9pt (default), 10pt, 11pt and 12pt
fontsize: 12pt

# Optional: Force one-column layout, default is two-column
one_column: true

# Optional: Enables lineno mode, but only if one_column mode is also true
#lineno: true

# Optional: Enable one-sided layout, default is two-sided
#one_sided: true

# Optional: Enable section numbering, default is unnumbered
#numbersections: true

# Optional: Specify the depth of section number, default is 5
#secnumdepth: 5

# Optional: Skip inserting final break between acknowledgements, default is false
skip_final_break: true

# Optional: Bibliography 
bibliography: pinp

# Optional: Enable a 'Draft' watermark on the document
watermark: false

# Customize footer, eg by referencing the vignette
# footer_contents: "YourPackage Vignette"

# Produce a pinp document
output: pinp::pinp

# Required: Vignette metadata for inclusion in a package.
vignette: >
  %\VignetteIndexEntry{YourPackage-vignetteentry}
  %\VignetteKeywords{YourPackage, r, anotherkeyword}
  %\VignettePackage{YourPackage}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path = "figures/", dpi=300, fig.height=6, fig.width=6)
# options(scipen = -2, digits = 2)
```

```{r read_chunks, cache=FALSE, echo=FALSE, eval=TRUE}
# source("code/simulation.R")
knitr::read_chunk("../scripts/simulation.R")
```

## Introduction

<!-- * Lake area estimates are important -->

Size is a critical factor determining the rate and occurence of specific lake processes such as carbon burial and sequestration \citep{delsontro_greenhouse_2018}. However, the relative contribution of small and large water bodies to overall nutrient and material cycling is unknown \citep{alexander2000effect}. One way to estimate their relative contributions is to combine information on per area processing rates with estimates of total area \citep{winslow_does_2015}.

<!-- * Lake area databases are censored -->

One of the challenges in estimating total lake area is that the distribution of individual lake areas spans a range of approximately 7 orders of magnitude whereby the largest lakes are so big they could otherwise be classified as inland seas (> $10^4$ $km^2$) while the smallest are barely larger than a regulation sized soccer field (> $10^{-3}$ $km^2$). Another important characteristic of lake area is that the area of the largest lakes is known exactly while the area of the smallest lakes is incomplete below a certain unknown threshold. This can occur either because lakes are too small to be resolved from remote sensing products or because of limited ground surveying effort. In this sense, lake databases are said to be censored at small lake areas because we know that small lakes exist but below a certain threshold we have limited knowledge of their exact areas \citep{hamiltonEstimationFractalDimension1992}.

<!-- * Lake area databases are truncated -->

Estimating total lake area from a sample of lakes requires a conceptual model of how lakes are formed (i.e. the data generating process). Typically, lake areas are treated as arising from a fractal generating process due to the fact that landform topography, which determines the placement of lakes, can itself be treated as a fractal generating process. Indeed, many other geomorphological phenomena that are dependent on landform topography such as coastline length are often well-described by fractal generating processes \citep{newman_power_2005}.

Another challenge in estimating total lake area is that the area distribution of the largest lakes likely follows a different data generating process than that of the smallest lakes. The reason that largest lakes likely follow a different data generating process is that they are constrained not by local landform topography but by the placement and arrangement of continents. As lakes become larger they have a greater probability of intersecting a continent edge and becoming an estuary or embayment rather than a lake \citep{goodchildLakesFractalSurfaces1988}. In this sense, lake databases are said to be truncated on large lakes because we know that large lakes are essentially fixed in space and cannot occur in any given location \citep{hamiltonEstimationFractalDimension1992}.

<!-- * Population estimates / Knowledge gap -->

From the preceding discussion it is clear that estimating total lake area requires a method of dealing both with the fact that lake databases are truncated at large lakes and censored at small lakes. Previous studies estimating global lake area have essentially not dealt with the first issue but instead modified their estimation process to limit its effect on the results (see methods section). They have dealt with the second issue by specifying an ad-hoc cutoff value below which empirically determined lake areas are discarded and subsequently back calculated. Hereafter, I refer to this strategy as the cutoff method. \textbf{In the present study, I explore the effects of using the typical frequentist approach for estimating total lake area (i.e. the cutoff method) via a simulation study. In addition, I demonstrate an approach to produce uncertainty estimates of total lake area in a Bayesian framework.} I evaluate the effects of the cutoff method on a simulated dataset because it is not sensitive to other potential confounding factors such as heterogeneity of survey effort or data precision.

## Methods

<!-- * Rationale for pareto model -->

Lake areas are typically treated as arising from a scale-invariant fractal generating process \citep{winslow_does_2015, downingGlobalAbundanceSize2006, mcdonald_regional_2012, goodchildLakesFractalSurfaces1988, hamiltonEstimationFractalDimension1992}. Essentially, this means that the number of lakes in one size class is proportional to the number of lakes in the preceding size class irrespective of their magnitudes. The numerical form describing such a process is a power-law function. One of the statistical tools often used to model data that follow a power-law function is the Pareto distribution which has a probability density function (pdf) of:

\begin{equation}
  \begin{aligned}
    pdf(A) = \alpha x_{min}^{\alpha}A^{-(\alpha+1)} \\
  \label{eqn:pareto_pdf}
  \end{aligned}
\end{equation}

where $A$ is lake area, $a$ controls the “shape” of the distribution and $x_{min}$ controls the “scale” of the distribution \citep{shaliziAdvancedDataAnalysis2017}. Lake area studies using the Pareto distribution do not typically use the pdf directly. Instead, they use the inverse (complementary) cumulative distribution function (ccdf) (i.e. quantile function):

\begin{equation}
  \begin{aligned}
    ccdf(A) = \frac{x_{min}}{(1-A)^{\frac{1}{\alpha}}}
  \label{eqn:pareto_ccdf}
  \end{aligned}
\end{equation}

The reason for using the ccdf is two-fold. First, it stabilizes model estimates in the lower tail of the distribution \citep{newman_power_2005}. This can be seen from the simulated data in Figure \ref{fig:pareto_demo} where the Pareto pdf contains a lot of noise in the tail but the ccdf appears smoothed. The smoothing of the tail is also desirable because it functions as a way of dealing with the truncated nature of lake databases. The second reason for using the ccdf is that it provides a computational shortcut for estimating the Pareto shape parameter $a$ because it is numerically equivalent to the slope of the ccdf in log-log space \citep{downingGlobalAbundanceSize2006}.

```{r pareto_demo, message=FALSE, results='hide', echo=FALSE, warning=FALSE, fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.", fig.height=2.5}

```

<!-- * Description of simulation study -->

For evaluation purposes, I generated a simulated dataset of 10,000 lake areas following the Pareto distribution using inverse transform sampling \citep{newman_power_2005}. Lakes in my simulated dataset have a minimum and maximum area of approximately 1 and 81000 $km^2$ respectively. This maximum was chosen to be approximately as large as Lake Superior but less than the Caspian Sea following \citet{lehnerDevelopmentValidationGlobal2004}. The "true" total area of these lakes is approximately `r as.character(round(sum(y), -4))` $km^2$. I simulated a censored lake dataset by excluding lakes smaller than $e^{1}$. This excludes (censors) approximately `r as.character(round(1 - length(y_censored) / length(y), 1) * 100)` % of the total dataset. I approximated the "true" lake area total by constructing the empirical distribution function (edf) of the data which approximates the underlying Pareto cdf \citep{newman_power_2005}. Then I used this estimate of the cdf slope to generate cdf estimates for the censored lakes. I combined these cdf estimates with the edf values from the known lakes before calculating the sum of the inverted distribution (Figure \ref{fig:predict_area}).

<!-- * Frequentist estimation -->

I estimated the Pareto shape parameter $a$ in a frequentist framework by calculating the the slope of the edf in log-log space using linear regression in `R` \citep{team2013r}. I evaluated uncertainty in both $a$ and total lake area in a Bayesian framework using Stan \citep{standevelopmentteamStanModelingLanguage2017}. Instead of computing on the edf (as in the frequentist case), I computed directly on the pdf with the following Stan model:

\singlespacing

```
data {
  int<lower=0> N; 
  real x[N];
} 
parameters {
  real<lower=0> alpha;
  real<lower=0> xmin;
} 
model { 
  real lpa[N];

  xmin ~ gamma(.001, .001);   
  alpha ~ gamma(.001, .001);

  for (i in 1:N) {  
    lpa[i] = pareto_lpdf(x[i] | xmin, alpha); 
  }

  target += sum(lpa);
}
```

\doublespacing
\FloatBarrier

I used vague gamma priors for both the $x_{min}$ and $a$ parameters following \citet{scollnikCompositeLognormalParetoModels2007}. I ran the model with four chains of 8,000 iterations and used the Stan defaults for burn-in and thinning which specify a burn-in of half the iterations and no thinning.

## Results

Visual inspection of the frequentist method of computing on the edf appeared to produce a reasonable density estimate for small censored lakes (Figure \ref{fig:predict_area}). In addition, estimates of total lake area are somewhat close to the "true" value (Table \ref{tab:pred_table}). However, uncertainty around the frequentist estimates is unrealistically small (Figure \ref{fig:frequentist_uncertainty}).

```{r predict_area, message=FALSE, results='hide', echo=FALSE, warning=FALSE, fig.height=2.5, fig.width=4.5, fig.cap="Censored lake area edf (solid line) and cdf estimate (dashed line)."}

```

```{r frequentist_uncertainty, message=FALSE, results='hide', echo=FALSE, warning=FALSE, fig.height=2.5, fig.width=4.5, fig.cap="Confidence interval of frequentist density predictions for censored data (the dashed portion in Figure 2. The confidence interval here has essentially no width."}

```

```{r bayesian_model, message=FALSE, results='hide', echo=FALSE, warning=FALSE, fig.height=2.5, fig.width=4.5, fig.cap="Median (black line) and central 95 percent interval estimates of alpha (red lines). Here the 'true' alpha is 0.9."}

```

```{r bayesian_area, message=FALSE, results='hide', echo=FALSE, warning=FALSE, fig.height=2.5, fig.width=4.5, fig.cap="Median (black line) and central 95 percent interval estimates of total lake area (red lines). Here the true total lake area is marked with a dashed vertical line."}

```

Instead of the essentially fixed $a$ and total lake area estimates produced by the frequentist approach, I found a substantial variability in both $a$ (95% CI: `r round(quantile(alphas, probs = c(0.025, .975)), 2)`) and total area using a Bayesian approach (Fig \ref{fig:bayesian_model}, \ref{fig:bayesian_area}). In particular, the Bayesian 95% credible intervals for both $a$ and total lake area encapsulate the true values (Fig \ref{fig:bayesian_model}, \ref{fig:bayesian_area}). Despite better uncertainty estimates using a Bayesian approach, both the frequentist and Bayesian approaches underestimated the true value of $a$ and total lake area.

```{r pred_table, message=FALSE, echo=FALSE, warning=FALSE}
kableExtra::kable_styling(
  knitr::kable(data.frame(`True Area (km2)` = formatC(total_empirical, 
                                         format = "e", digits = 3), 
                          `Estimated Area (km2)` = formatC(total_predicted, 
                                         format = "e", digits = 3), 
                          check.names = FALSE), 
               format = "latex", booktabs = TRUE, align = c("cc"),
               caption = "'True' total lake area from the uncensored edf and 
estimated lake area from a combination of the censored edf and the estimated cdf."), 
  latex_options = "hold_position")
```

## Discussion

I have shown that the typical frequentist cutoff method produces reasonable estimates of the density of small censored lakes but that it does not capture uncertainty in total lake area (Table \ref{tab:pred_table}, Figure \ref{fig:frequentist_uncertainty}). Furthermore, I have shown that models fit using a Bayesian approach indicate substantial uncertainty in both total lake area and the underlying Pareto shape parameter $a$ used to derive these estimates (Figure \ref{fig:bayesian_model}, \ref{fig:bayesian_area}).

Although the 95% credible interval of the Bayesian total lake area estimates encapsulate the true total lake area, the median value underestimates the true total lake area (Figure \ref{fig:bayesian_area}). It is likely that such underestimation, would increase with a greater proportion of censoring. This may explain the steady increase in estimates of global lake area through time from approximately 3 to 5 million $km^{2}$ as lake area databases have improved their accuracy \citep{lehnerDevelopmentValidationGlobal2004, downingGlobalAbundanceSize2006, verpoorterGlobalInventoryLakes2014}. Future work on estimating global lake area should consider implementing a sensitivity analysis looking at the response of total area estimates to variation in the degree of censoring.

Future investigators should be mindful that is difficult to confirm whether or not any particular dataset following a fractal generating or Pareto process without strong prior knowledge. Indeed data following many heavy-tailed distributions such as the lognormal or negative exponential can appear to be equivalent to power-law distributions such as the Pareto \citep{clausetPowerlawDistributionsEmpirical2009}. The results of the present study confirm that an apparent change in the shape of the lower tail of a lake area distribution does not necessarily indicate a change in the data generating process. Note that the simulation dataset analyzed herein _is_ a truly fractal generated dataset with a homogeneous data generating process yet a cursory look would seem to indicate a change of data-generating process in the lower tail (Figure \ref{fig:predict_area}).

In addition to a sensitivity analysis of censoring, future work should consider more complex models that treat lake areas as a mixture of a Pareto distribution for small lakes and either a negative exponential or lognormal distribution for large lakes. Such an approach has been demonstrated by \citet{bonabeau_scaling_1999-1} and \citet{scollnikCompositeLognormalParetoModels2007}. Both studies have shown that the point at which the distribution mixtures converge can provide valuable inference. In the case of lakes, such a convergence point may indicate a change in the data generating process (i.e. the point at which lake area are controlled by continent placement rather than fractal landscape morphology \citep{goodchildLakesFractalSurfaces1988, hamiltonEstimationFractalDimension1992}.

### 2019-03-05 scripts/simulation.R

@@ -10,18 +10,18 @@ qpareto <- function(u, a=0.9, b=1) b/(1-u)^(1/a)
 rpareto <- function(n, a=0.9, b=1, max = Inf){
   res <- qpareto(runif(n), a, b)
   res <- res[res < max]
-  
+
   while(length(res) < n){
     res <- c(res, qpareto(runif(n - length(res)), a, b))
   }
   res
-} 
+}
 # (dt <- sample(1:10, 12, replace = TRUE) %>% .[order(.)])
 # (cf <- cumulative_freq(dt))
 cumulative_freq <- function(x){
   l_n    <- cumsum(rev(table(x)))
   l_area <- rev(as.numeric(names(table(x))))
-  
+
   data.frame(area = l_area, number = l_n)
 }
 
@@ -32,25 +32,28 @@ inv_cumulative_freq <- function(cf){
 }
 
 # simulate random pareto draws
-# y <- rpareto(10000, max = 81935.7) # cap at the area of Lake Superior
-# saveRDS(y, "project/paper/data/y.rds")
-y <- readRDS("data/y.rds")
+if(!file.exists("../data/y.rds")){
+  y <- rpareto(10000, max = 81935.7) # cap at the area of Lake Superior
+  saveRDS(y, "../data/y.rds")
+}
+#
+y <- readRDS("../data/y.rds")
 
 plot_grid(
   # individual binning
-  hist(log(y), plot = FALSE, n = 100) %>% 
+  hist(log(y), plot = FALSE, n = 100) %>%
   {data.frame(x = .$breaks[-1], samples = log(.$counts))} %>%
     ggplot + geom_line(aes(x, samples)) +
     theme_minimal(),
   # log binning
-  # hist(y, plot = FALSE, breaks = 10^(seq(from = 0, to = 6, by = 0.6))) %>% 
+  # hist(y, plot = FALSE, breaks = 10^(seq(from = 0, to = 6, by = 0.6))) %>%
   # {data.frame(x = .$breaks[-1], samples = .$counts + 0.01)} %>%
   #   ggplot + geom_line(aes(x, samples)) +
   #   theme_minimal() + scale_x_log10() + scale_y_log10(),
   # cumulative binning
-  cumulative_freq(y) %>% 
+  cumulative_freq(y) %>%
     ggplot + geom_line(aes(area, number)) +
-    theme_minimal() + scale_x_log10() + scale_y_log10() + 
+    theme_minimal() + scale_x_log10() + scale_y_log10() +
[…]

### 2021-05-15 manuscript/manuscript.Rmd
msu->UW

@@
 ---
-title: Estimating Uncertainty in Global Lake Area
+title: Quantifying uncertainty in Pareto lake area estimates
 
@@
   - code: a
-    address: Department of Fisheries and Wildlife, Michigan State University, East Lansing, MI, USA
+    address: Center for Limnology, University of Wisconsin-Madison, Madison, WI, USA
 

### 2021-05-15 manuscript/template.tex
stash rmd to tex conversion

--- 2021-05-15 manuscript/manuscript.Rmd
@@
----
-title: Quantifying uncertainty in Pareto lake area estimates
+\documentclass{article}
 
-author:
-  - name: Joseph Stachelek
-    affiliation: a
-address:
-  - code: a
-    address: Center for Limnology, University of Wisconsin-Madison, Madison, WI, USA
+\usepackage{arxiv}
 
-lead_author_surname: Stachelek
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+%\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+\usepackage{lipsum}		% Can be removed after putting your text content
+\usepackage{graphicx}
+\usepackage{natbib}
+\usepackage{doi}
+\usepackage{listings}
 
-# Place DOI URL or CRAN Package URL here
-# doi: "https://cran.r-project.org/package=YourPackage"
+\title{Quantifying uncertainty in Pareto lake area estimates}
 
-abstract: | 
-  Size is a critical factor determining the rate and occurrence of specific lake processes such as carbon sequestration.
-Unfortunately, the exact lake size-abundance distribution is unknown in part because we do not have a complete census of all lakes.
-More specifically, as lakes become smaller they are less likely to be included in lake databases either because they are too small to be resolved from remote sensing products or because of limited ground surveying effort.
-The exclusion of small lakes relative to large lakes is known as censoring.
+%\date{September 9, 1985}	% Here you can change the date presented in the paper title
+%\date{} 					% Or removing it
 
-  The present study explores one potential shortcoming of previous approaches at estimating global lake area and dealing with censoring.
-Namely, that the typical frequentist curve fitting and ad-hoc cutoff determination strategy (visual inspection to determine a likely censoring point) taken by previous studies yields an over-exact lake area estimate with no reported uncertainty.
+\author{ \href{https://orcid.org/0000-0000-0000-0000}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}J.~Stachelek}\thanks{\href{https://jsta.rbind.io}{https://jsta.rbind.io}} \\
+	Center for Limnology\\
+	University of Wisconsin-Madison\\
+	Madison, WI 53706 \\
+	\texttt{stachelek@wisc.edu} \\
+}
+
+% Uncomment to remove the date
+%\date{}
+
+% Uncomment to override  the `A preprint' in the header
+%\renewcommand{\headeright}{Technical Report}
+%\renewcommand{\undertitle}{Technical Report}
+\renewcommand{\shorttitle}{Pareto lake area uncertainty}
+
+%%% Add PDF metadata to help others organize their library
+%%% Once the PDF is generated, you can check the metadata with
+%%% $ pdfinfo template.pdf
+\hypersetup{
+pdftitle={Quantifying uncertainty in Pareto lake area estimates},
+pdfsubject={q-bio.NC, q-bio.QM},
+pdfauthor={J.~Stachelek}
+}
+
+\begin{document}
+\maketitle
+
+\begin{abstract}
+	Size is a critical factor determining the rate and occurrence of specific lake processes such as carbon sequestration relative to greenhouse gas emissions.
+Upscaling estimates of such processes at broad spatial scales requires the use of lake size-abundance distributions rather than empirical measurements of area because we do not have a complete census of all lakes.
+The reason for this is that as lakes become smaller they are more likely to not be included in lake databases either because they are too small to be resolved from remote sensing products or because of limited ground surveying effort (i.e.
+"censoring" of small lakes relative to large lakes).
+
+  The present study explores one potential shortcoming of prior approaches estimating global lake area using lake size-abundance distributions.
+Namely, these prior approaches rely on frequentist curve fitting techniques combined with an ad-hoc cutoff determination strategy (visual inspection to determine a likely censoring point) taken by previous studies yields an over-exact lake area estimate with no reported uncertainty.
 I address this shortcoming by fitting models in a Bayesian framework where each parameter contributes uncertainty to model estimates.
@@
 Ultimately, this may explain the fact that total lake area estimates have increased through time as the resolution of lake databases has improved.
+\end{abstract}
 
-# Optional: Acknowledgements
-# acknowledgements: |
+% keywords can be removed
+%\keywords{lake area \and pareto \and limnology}
 
-# Optional: One or more keywords
-# keywords:
-#   - one
-#   - two
-#   - optional
-#   - keywords
-#   - here
-
-# Font size of the document, values of 9pt (default), 10pt, 11pt and 12pt
-fontsize: 12pt
-
-# Optional: Force one-column layout, default is two-column
-one_column: true
-
-# Optional: Enables lineno mode, but only if one_column mode is also true
-#lineno: true
-
-# Optional: Enable one-sided layout, default is two-sided
-#one_sided: true
-
-# Optional: Enable section numbering, default is unnumbered
-#numbersections: true
-
-# Optional: Specify the depth of section number, default is 5
-#secnumdepth: 5
-
-# Optional: Skip inserting final break between acknowledgements, default is false
-skip_final_break: true
-
-# Optional: Bibliography 
-bibliography: pinp
-
-# Optional: Enable a 'Draft' watermark on the document
-watermark: false
-
-# Customize footer, eg by referencing the vignette
-# footer_contents: "YourPackage Vignette"
-
-# Produce a pinp document
-output: pinp::pinp
-
-# Required: Vignette metadata for inclusion in a package.
-vignette: >
-  %\VignetteIndexEntry{YourPackage-vignetteentry}
-  %\VignetteKeywords{YourPackage, r, anotherkeyword}
-  %\VignettePackage{YourPackage}
-  %\VignetteEngine{knitr::rmarkdown}
----
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, fig.path = "figures/", dpi=300, fig.height=6, fig.width=6)
-# options(scipen = -2, digits = 2)
-```
-
-```{r read_chunks, cache=FALSE, echo=FALSE, eval=TRUE}
-# source("code/simulation.R")
-knitr::read_chunk("../scripts/simulation.R")
-```
-
-## Introduction
-
-<!-- * Lake area estimates are important -->
-
-Size is a critical factor determining the rate and occurence of specific lake processes such as carbon burial and sequestration \citep{delsontro_greenhouse_2018}.
+\section{Introduction}
+Size is a critical factor determining the rate and occurence of specific lake processes such as carbon burial and sequestration \citep{delsontroGreenhouseGasEmissions2018, kellerGlobalCarbonBudget2021}.
 However, the relative contribution of small and large water bodies to overall nutrient and material cycling is unknown \citep{alexander2000effect}.
-One way to estimate their relative contributions is to combine information on per area processing rates with estimates of total area \citep{winslow_does_2015}.
-
-<!-- * Lake area databases are censored -->
+One way to estimate their relative contributions is to combine information on per area processing rates with estimates of total area \citep{winslowDoesLakeSize2015}.
 
@@
 In this sense, lake databases are said to be censored at small lake areas because we know that small lakes exist but below a certain threshold we have limited knowledge of their exact areas \citep{hamiltonEstimationFractalDimension1992}.
-
-<!-- * Lake area databases are truncated -->
 
@@
 
-<!-- * Population estimates / Knowledge gap -->
-
 From the preceding discussion it is clear that estimating total lake area requires a method of dealing both with the fact that lake databases are truncated at large lakes and censored at small lakes.
@@
 Hereafter, I refer to this strategy as the cutoff method.
-\textbf{In the present study, I explore the effects of using the typical frequentist approach for estimating total lake area (i.e.
+In the present study, I explore the effects of using the typical frequentist approach for estimating total lake area (i.e.
 the cutoff method) via a simulation study.
-In addition, I demonstrate an approach to produce uncertainty estimates of total lake area in a Bayesian framework.} I evaluate the effects of the cutoff method on a simulated dataset because it is not sensitive to other potential confounding factors such as heterogeneity of survey effort or data precision.
+In addition, I demonstrate an approach to produce uncertainty estimates of total lake area in a Bayesian framework.
+I evaluate the effects of the cutoff method on a simulated dataset because it is not sensitive to other potential confounding factors such as heterogeneity of survey effort or data precision.
 
-## Methods
+\section{Methods}
 
-<!-- * Rationale for pareto model -->
-
-Lake areas are typically treated as arising from a scale-invariant fractal generating process \citep{winslow_does_2015, downingGlobalAbundanceSize2006, mcdonald_regional_2012, goodchildLakesFractalSurfaces1988, hamiltonEstimationFractalDimension1992}.
+Lake areas are typically treated as arising from a scale-invariant fractal generating process \citep{winslowDoesLakeSize2015, downingGlobalAbundanceSize2006, mcdonald_regional_2012, goodchildLakesFractalSurfaces1988, hamiltonEstimationFractalDimension1992}.
 Essentially, this means that the number of lakes in one size class is proportional to the number of lakes in the preceding size class irrespective of their magnitudes.
@@
 \begin{equation}
-  \begin{aligned}
-    pdf(A) = \alpha x_{min}^{\alpha}A^{-(\alpha+1)} \\
-  \label{eqn:pareto_pdf}
-  \end{aligned}
-\end{equation}
-
-where $A$ is lake area, $a$ controls the “shape” of the distribution and $x_{min}$ controls the “scale” of the distribution \citep{shaliziAdvancedDataAnalysis2017}.
-Lake area studies using the Pareto distribution do not typically use the pdf directly.
-Instead, they use the inverse (complementary) cumulative distribution function (ccdf) (i.e.
-quantile function):
-
-\begin{equation}
-  \begin{aligned}
     ccdf(A) = \frac{x_{min}}{(1-A)^{\frac{1}{\alpha}}}
   \label{eqn:pareto_ccdf}
-  \end{aligned}
 \end{equation}
@@
 
-```{r pareto_demo, message=FALSE, results='hide', echo=FALSE, warning=FALSE, fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.", fig.height=2.5}
-
-```
-
-<!-- * Description of simulation study -->
+\begin{figure}
+	\centering
+	\includegraphics{figures/pareto_demo-1}
+	\caption{Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.}
+	\label{fig:pareto_demo}
+\end{figure}
 
@@
 This maximum was chosen to be approximately as large as Lake Superior but less than the Caspian Sea following \citet{lehnerDevelopmentValidationGlobal2004}.
-The "true" total area of these lakes is approximately `r as.character(round(sum(y), -4))` $km^2$.
+The "true" total area of these lakes is approximately 230000 $km^2$.
 I simulated a censored lake dataset by excluding lakes smaller than $e^{1}$.
-This excludes (censors) approximately `r as.character(round(1 - length(y_censored) / length(y), 1) * 100)` % of the total dataset.
+This excludes (censors) approximately 60\% of the total dataset.
 I approximated the "true" lake area total by constructing the empirical distribution function (edf) of the data which approximates the underlying Pareto cdf \citep{newman_power_2005}.
@@
 
-<!-- * Frequentist estimation -->
-
-I estimated the Pareto shape parameter $a$ in a frequentist framework by calculating the the slope of the edf in log-log space using linear regression in `R` \citep{team2013r}.
+I estimated the Pareto shape parameter $a$ in a frequentist framework by calculating the the slope of the edf in log-log space using linear regression in \texttt{R} \citep{team2013r}.
 I evaluated uncertainty in both $a$ and total lake area in a Bayesian framework using Stan \citep{standevelopmentteamStanModelingLanguage2017}.
@@
 
-\singlespacing
-
-```
+\begin{minipage}[c]{0.95\textwidth}
+\begin{lstlisting}
 data {
@@
 }
-```
-
-\doublespacing
-\FloatBarrier
+\end{lstlisting}
+\end{minipage}
 
@@
 
-## Results
+\section{Results}
 
@@
 
-```{r predict_area, message=FALSE, results='hide', echo=FALSE, warning=FALSE, fig.height=2.5, fig.width=4.5, fig.cap="Censored lake area edf (solid line) and cdf estimate (dashed line)."}
+\begin{figure}
+	\centering
+	\includegraphics{figures/predict_area-1}
+	\caption{Censored lake area edf (solid line) and cdf estimate (dashed line).}
+	\label{fig:predict_area}
+\end{figure}
 
-```
+\begin{figure}
+	\centering
+	\includegraphics{figures/frequentist_uncertainty-1}
+	\caption{Confidence interval of frequentist density predictions for censored data (the dashed portion in Figure 2. The confidence interval here has essentially no width.}
+	\label{fig:frequentist_uncertainty}
+\end{figure}
 
-```{r frequentist_uncertainty, message=FALSE, results='hide', echo=FALSE, warning=FALSE, fig.height=2.5, fig.width=4.5, fig.cap="Confidence interval of frequentist density predictions for censored data (the dashed portion in Figure 2.
-The confidence interval here has essentially no width."}
+\begin{figure}
+	\centering
+	\includegraphics{figures/bayesian_model-1}
+	\caption{Median (black line) and central 95 percent interval estimates of alpha (red lines). Here the 'true' alpha is 0.9.}
+	\label{fig:bayesian_model}
+\end{figure}
 
-```
+\begin{figure}
+	\centering
+	\includegraphics{figures/bayesian_area-1}
+	\caption{Median (black line) and central 95 percent interval estimates of total lake area (red lines). Here the true total lake area is marked with a dashed vertical line.}
+	\label{fig:bayesian_area}
+\end{figure}
 
-```{r bayesian_model, message=FALSE, results='hide', echo=FALSE, warning=FALSE, fig.height=2.5, fig.width=4.5, fig.cap="Median (black line) and central 95 percent interval estimates of alpha (red lines).
-Here the 'true' alpha is 0.9."}
-
-```
-
-```{r bayesian_area, message=FALSE, results='hide', echo=FALSE, warning=FALSE, fig.height=2.5, fig.width=4.5, fig.cap="Median (black line) and central 95 percent interval estimates of total lake area (red lines).
-Here the true total lake area is marked with a dashed vertical line."}
-
-```
-
-Instead of the essentially fixed $a$ and total lake area estimates produced by the frequentist approach, I found a substantial variability in both $a$ (95% CI: `r round(quantile(alphas, probs = c(0.025, .975)), 2)`) and total area using a Bayesian approach (Fig \ref{fig:bayesian_model}, \ref{fig:bayesian_area}).
-In particular, the Bayesian 95% credible intervals for both $a$ and total lake area encapsulate the true values (Fig \ref{fig:bayesian_model}, \ref{fig:bayesian_area}).
+Instead of the essentially fixed $a$ and total lake area estimates produced by the frequentist approach, I found a substantial variability in both $a$ (95\% CI: 0.86, 0.92) and total area using a Bayesian approach (Fig \ref{fig:bayesian_model}, \ref{fig:bayesian_area}).
+In particular, the Bayesian 95\% credible intervals for both $a$ and total lake area encapsulate the true values (Fig \ref{fig:bayesian_model}, \ref{fig:bayesian_area}).
 Despite better uncertainty estimates using a Bayesian approach, both the frequentist and Bayesian approaches underestimated the true value of $a$ and total lake area.
 
-```{r pred_table, message=FALSE, echo=FALSE, warning=FALSE}
-kableExtra::kable_styling(
-  knitr::kable(data.frame(`True Area (km2)` = formatC(total_empirical, 
-                                         format = "e", digits = 3), 
-                          `Estimated Area (km2)` = formatC(total_predicted, 
-                                         format = "e", digits = 3), 
-                          check.names = FALSE), 
-               format = "latex", booktabs = TRUE, align = c("cc"),
-               caption = "'True' total lake area from the uncensored edf and 
-estimated lake area from a combination of the censored edf and the estimated cdf."), 
-  latex_options = "hold_position")
-```
+\begin{table}
+	\caption{'True' total lake area from the uncensored edf and estimated lake area from a combination of the censored edf and the estimated cdf.}
+	\centering
+	\begin{tabular}{ll}
+	\toprule
+		True Area ($km^2$) & Estimated Area ($km^2$) \\
+	\midrule
+		2.271e+05 & 2.263e+05 \\
+	\bottomrule
+	\end{tabular}
+	\label{tab:pred_table}
+\end{table}
 
-## Discussion
+\section{Discussion}
 
@@
 
-Although the 95% credible interval of the Bayesian total lake area estimates encapsulate the true total lake area, the median value underestimates the true total lake area (Figure \ref{fig:bayesian_area}).
+Although the 95\% credible interval of the Bayesian total lake area estimates encapsulate the true total lake area, the median value underestimates the true total lake area (Figure \ref{fig:bayesian_area}).
 It is likely that such underestimation, would increase with a greater proportion of censoring.
@@
 The results of the present study confirm that an apparent change in the shape of the lower tail of a lake area distribution does not necessarily indicate a change in the data generating process.
-Note that the simulation dataset analyzed herein _is_ a truly fractal generated dataset with a homogeneous data generating process yet a cursory look would seem to indicate a change of data-generating process in the lower tail (Figure \ref{fig:predict_area}).
+Note that the simulation dataset analyzed herein \emph{is} a truly fractal generated dataset with a homogeneous data generating process yet a cursory look would seem to indicate a change of data-generating process in the lower tail (Figure \ref{fig:predict_area}).
 
@@
 the point at which lake area are controlled by continent placement rather than fractal landscape morphology \citep{goodchildLakesFractalSurfaces1988, hamiltonEstimationFractalDimension1992}.
+
+\bibliographystyle{unsrtnat}
+\bibliography{pareto-lakes}
+
+\end{document}

### 2021-05-16 manuscript/manuscript.tex
stash text edit cleanup

--- 2021-05-15 manuscript/template.tex
@@
 
-\author{ \href{https://orcid.org/0000-0000-0000-0000}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}J.~Stachelek}\thanks{\href{https://jsta.rbind.io}{https://jsta.rbind.io}} \\
+\author{ \href{https://orcid.org/0000-0002-5924-2464}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}J.~Stachelek}\thanks{\href{https://jsta.rbind.io}{https://jsta.rbind.io}} \\
 	Center for Limnology\\
@@
 \begin{abstract}
-	Size is a critical factor determining the rate and occurrence of specific lake processes such as carbon sequestration relative to greenhouse gas emissions.
-Upscaling estimates of such processes at broad spatial scales requires the use of lake size-abundance distributions rather than empirical measurements of area because we do not have a complete census of all lakes.
-The reason for this is that as lakes become smaller they are more likely to not be included in lake databases either because they are too small to be resolved from remote sensing products or because of limited ground surveying effort (i.e.
+	Size is a critical factor determining the rate and occurrence of specific lake processes such as carbon sequestration and greenhouse gas emissions.
+Because we do not have a complete census of all lakes, upscaling estimates of such processes at broad spatial scales requires the use of lake size-abundance distributions rather than empirical measurements of area.
+Existing lake census efforts are incomplete such that as lakes become smaller they are more likely to by omitted either because they are too small to be resolved from remote sensing products or because of limited ground surveying effort (i.e.
 "censoring" of small lakes relative to large lakes).
@@
 The present study explores one potential shortcoming of prior approaches estimating global lake area using lake size-abundance distributions.
-Namely, these prior approaches rely on frequentist curve fitting techniques combined with an ad-hoc cutoff determination strategy (visual inspection to determine a likely censoring point) taken by previous studies yields an over-exact lake area estimate with no reported uncertainty.
+Namely, these prior approaches rely on frequentist curve fitting techniques combined with an ad-hoc cutoff determination strategy (visual inspection to determine a likely censoring point).
+This yields an over-exact lake area estimate that is overwhelmingly reported with no uncertainty bounds.
 I address this shortcoming by fitting models in a Bayesian framework where each parameter contributes uncertainty to model estimates.
@@
 
-% keywords can be removed
-%\keywords{lake area \and pareto \and limnology}
+\section{Introduction}
+Size is a critical factor determining the contribution of lakes to global biogeochemical cycling and existing evidence suggests that lakes are key components of overall carbon burial, sequestration, and greenhouse gas emissions \citep{delsontroGreenhouseGasEmissions2018, kellerGlobalCarbonBudget2021}.
+The calculation of such fluxes requires two terms 1) an areal flux rate and 2) the distribution of total lake area.
+As with all estimates at broad spatial scales, uncertainty estimates are of great interest both because they allow for intercomparisons among similar lake studies and because they allow for comparisons between lakes and other ecosystems.
 
-\section{Introduction}
-Size is a critical factor determining the rate and occurence of specific lake processes such as carbon burial and sequestration \citep{delsontroGreenhouseGasEmissions2018, kellerGlobalCarbonBudget2021}.
-However, the relative contribution of small and large water bodies to overall nutrient and material cycling is unknown \citep{alexander2000effect}.
-One way to estimate their relative contributions is to combine information on per area processing rates with estimates of total area \citep{winslowDoesLakeSize2015}.
-
-One of the challenges in estimating total lake area is that the distribution of individual lake areas spans a range of approximately 7 orders of magnitude whereby the largest lakes are so big they could otherwise be classified as inland seas (> $10^4$ $km^2$) while the smallest are barely larger than a regulation sized soccer field (> $10^{-3}$ $km^2$).
-Another important characteristic of lake area is that the area of the largest lakes is known exactly while the area of the smallest lakes is incomplete below a certain unknown threshold.
+The techniques typically used to determine uncertainty in areal flux rates and the techniques typically used to determine uncertainty in total lake area distributions differ greatly.
+In the former case, uncertainties can typically be calculated in straightforward manner from literature compilations \citep{delsontroGreenhouseGasEmissions2018, kellerGlobalCarbonBudget2021}.
+In the latter case, uncertainty in total lake area distributions is challenged by the fact that no existing database is a complete census of all lakes \citep{messagerEstimatingVolumeAge2016} and the distribution of individual lake areas spans a range of approximately 7 orders of magnitude.
+The largest lakes are so big they could otherwise be classified as inland seas (> $10^4$ $km^2$) while the smallest are barely larger than a regulation sized soccer field (> $10^{-3}$ $km^2$).
+A consequence of these challenges is that uncertainty varies within the distribution as a function of lake size.
+The area of the largest lakes is known exactly while the area of the smallest lakes is incomplete below a certain unknown threshold.
 This can occur either because lakes are too small to be resolved from remote sensing products or because of limited ground surveying effort.
-In this sense, lake databases are said to be censored at small lake areas because we know that small lakes exist but below a certain threshold we have limited knowledge of their exact areas \citep{hamiltonEstimationFractalDimension1992}.
+The omission or "censoring" of small lakes occurs because we know that small lakes exist but below a certain threshold we have limited knowledge of their exact areas \citep{hamiltonEstimationFractalDimension1992}.
 
@@
 Indeed, many other geomorphological phenomena that are dependent on landform topography such as coastline length are often well-described by fractal generating processes \citep{newman_power_2005}.
+A challenge in modelling such data generating processes for lake areas is large lakes likely follow a different data generating process than that of the smallest lakes.
+Whereas small lakes are constrained by landform topography, large lakes are essentially unconstrained by local landform topography and are instead constrained by the placement and arrangement of continents \citep{goodchildLakesFractalSurfaces1988}.
+As large lakes become even larger, they have a greater probability of intersecting a continent edge and becoming an estuary or embayment rather than a lake.
+Consquently, lake databases are said to be truncated on large lakes because we know that large lakes are essentially fixed in space and cannot occur in any given location \citep{hamiltonEstimationFractalDimension1992}.
 
-Another challenge in estimating total lake area is that the area distribution of the largest lakes likely follows a different data generating process than that of the smallest lakes.
-The reason that largest lakes likely follow a different data generating process is that they are constrained not by local landform topography but by the placement and arrangement of continents.
-As lakes become larger they have a greater probability of intersecting a continent edge and becoming an estuary or embayment rather than a lake \citep{goodchildLakesFractalSurfaces1988}.
-In this sense, lake databases are said to be truncated on large lakes because we know that large lakes are essentially fixed in space and cannot occur in any given location \citep{hamiltonEstimationFractalDimension1992}.
-
-From the preceding discussion it is clear that estimating total lake area requires a method of dealing both with the fact that lake databases are truncated at large lakes and censored at small lakes.
+From the preceding discussion it is clear that estimating total lake area requires a method of dealing both with the fact that lake databases are 1) truncated at large lakes and 2) censored at small lakes.
 Previous studies estimating global lake area have essentially not dealt with the first issue but instead modified their estimation process to limit its effect on the results (see methods section).
@@
 One of the statistical tools often used to model data that follow a power-law function is the Pareto distribution which has a probability density function (pdf) of:
+
+\begin{equation}
+    pdf(A) = \alpha x_{min}^{\alpha}A^{-(\alpha+1)} \\
+  \label{eqn:pareto_pdf}
+\end{equation}
+
+where $A$ is lake area, $a$ controls the “shape” of the distribution and $x_{min}$ controls the “scale” of the distribution \citep{shaliziAdvancedDataAnalysis2017}.
+Lake area studies using the Pareto distribution do not typically use the pdf directly.
+Instead, they use the inverse (complementary) cumulative distribution function (ccdf) (i.e.
+quantile function):
 
@@
 
-I used vague gamma priors for both the $x_{min}$ and $a$ parameters following \citet{scollnikCompositeLognormalParetoModels2007}.
+I used uninformative gamma priors for both the $x_{min}$ and $a$ parameters following \citet{scollnikCompositeLognormalParetoModels2007}.
 I ran the model with four chains of 8,000 iterations and used the Stan defaults for burn-in and thinning which specify a burn-in of half the iterations and no thinning.
@@
 	\centering
-	\begin{tabular}{ll}
+	\begin{tabular}{cc}
 	\toprule
@@
 
-Future investigators should be mindful that is difficult to confirm whether or not any particular dataset following a fractal generating or Pareto process without strong prior knowledge.
+A caveat of the present study is that it is difficult to confirm whether or not any particular dataset follows a fractal generating or Pareto process without strong prior knowledge.
 Indeed data following many heavy-tailed distributions such as the lognormal or negative exponential can appear to be equivalent to power-law distributions such as the Pareto \citep{clausetPowerlawDistributionsEmpirical2009}.
@@
 
-In addition to a sensitivity analysis of censoring, future work should consider more complex models that treat lake areas as a mixture of a Pareto distribution for small lakes and either a negative exponential or lognormal distribution for large lakes.
+In addition to a sensitivity analysis of censoring, future work might consider more complex models that treat lake areas as a mixture of a Pareto distribution for small lakes and either a negative exponential or lognormal distribution for large lakes.
 Such an approach has been demonstrated by \citet{bonabeau_scaling_1999-1} and \citet{scollnikCompositeLognormalParetoModels2007}.
-Both studies have shown that the point at which the distribution mixtures converge can provide valuable inference.
+Both studies show that the point at which the distribution mixtures converge can provide valuable inference.
 In the case of lakes, such a convergence point may indicate a change in the data generating process (i.e.
@@
 
-\bibliographystyle{unsrtnat}
+\bibliographystyle{jsta}
 \bibliography{pareto-lakes}

### 2021-05-16 scripts/00_get_iris.R
rm template junk, stash figure pdfs

@@ -1,6 +1,0 @@
-
-iris_path <- "data/iris.csv"
-# if(!file.exists(iris_path)){
-data("iris")
-write.csv(iris, iris_path, row.names = FALSE)
-# }

### 2021-05-16 scripts/01.5_loaddata.R

@@ -1,1 +1,0 @@
-iris <- read.csv("../data/iris.csv", stringsAsFactors = FALSE)

### 2021-05-16 scripts/01_prepdata.R

@@ -1,7 +1,0 @@
-# place any data preparation steps here
-
-library(tidyr)
-
-iris <- read.csv("data/iris.csv", stringsAsFactors = FALSE)
-iris_tidy_path <- "data/iris_tidy.csv"
-iris_tidy      <- tidyr::gather(iris, "variable", "value", -Species)

### 2021-05-16 scripts/02_modelling.R

@@ -1,10 +1,0 @@
-# setwd("scripts")
-source("99_utils.R")
-source("01_prepdata.R")
-
-# ---- load_packages ----
-library(broom)
-
-# ---- iris_lm ----
-fit <- lm(Sepal.Length ~ Petal.Length, data = iris)
-broom::tidy(fit)

### 2021-05-16 scripts/03_viz.R

@@ -1,5 +1,0 @@
-# setwd("scripts")
-# source("99_utils.R")
-# source("01_prepdata.R")
-
-# optionally construct plots here instead of in Rmd files

### 2021-05-16 scripts/04_tables.R

@@ -1,5 +1,0 @@
-# setwd("scripts")
-source("99_utils.R")
-# source("01_prepdata.R")
-
-# optionally construct tables here instead of in Rmd files

### 2021-05-16 scripts/99_utils.R

@@ -1,23 +1,0 @@
-
-# ---- source_utils ----
-
-library(ggplot2)
-
-theme_pred <- function(){
-  theme(legend.position = "na",
-        axis.text = element_text(size = 10),
-        axis.title = element_text(size = 12),
-        plot.title = element_text(size = 10, face = "bold", color = "black", hjust = 0))
-}
-
-signif_star <- function(x){
-  if(!is.na(x)){
-    if(x){
-      "*"
-    }else{
-      ""
-    }
-  }else{
-    ""
-  }
-}

### 2021-05-16 scripts/simulation.R
make build works

@@ -1,4 +1,5 @@
 # ---- pareto_demo ----
+# setwd("scripts")
 library(dplyr)
 library(magrittr)
 library(ggplot2)
@@ -39,7 +40,7 @@ if(!file.exists("../data/y.rds")){
 #
 y <- readRDS("../data/y.rds")
 
-plot_grid(
+pareto_demo <- plot_grid(
   # individual binning
   hist(log(y), plot = FALSE, n = 100) %>%
   {data.frame(x = .$breaks[-1], samples = log(.$counts))} %>%
@@ -56,6 +57,8 @@ plot_grid(
     theme_minimal() + scale_x_log10() + scale_y_log10() +
     ylab("samples with value > x") + xlab("x"),
   nrow = 1)
+ggsave("../manuscript/figures/pareto_demo.pdf", pareto_demo,
+       width = 5.93, height = 2.33)
 
 # remove lakes below censor threshold ####
 # hist(log(y))
@@ -96,9 +99,11 @@ cf_extra$number_upper  <- exp(cf_extra$upper + max(log(cf$number)))
 
 res         <- dplyr::bind_rows(cf_extra, cf)
 
-ggplot(data = res) +
+predict_area <- ggplot(data = res) +
   geom_line(aes(x = area, y = density, linetype = type)) +
   scale_x_log10() + theme(legend.title = element_blank())
+ggsave("../manuscript/figures/predict_area.pdf", predict_area,
+       width = 4.28, height = 2.33)
 
 # back-out an estimate of total area
 total_predicted <- sum(inv_cumulative_freq(res))
@@ -115,9 +120,11 @@ stack_preds <- data.frame(preds = stack_preds,
 #   res$number_upper[(nrow(cf_extra) + 1):nrow(res)] <-
 #   res$number[(nrow(cf_extra) + 1):nrow(res)]
 
-ggplot() + geom_line(data = stack_preds, aes(x = area, y = preds, color = type)) +
+frequentist_uncertainty <- ggplot() +
+  geom_line(data = stack_preds, aes(x = area, y = preds, color = type)) +
   scale_x_log10() + xlab("density") + labs(color = "Confidence \n Interval")
-
+ggsave("../manuscript/figures/frequentist_uncertainty.pdf", frequentist_uncertainty,
+       width = 5.93, height = 2.33)
 
 # ---- bayesian_model ----
[…]

### 2021-05-16 scripts/simulation.R
list of figures in makefile build

@@ -57,7 +57,7 @@ pareto_demo <- plot_grid(
     theme_minimal() + scale_x_log10() + scale_y_log10() +
     ylab("samples with value > x") + xlab("x"),
   nrow = 1)
-ggsave("../manuscript/figures/pareto_demo.pdf", pareto_demo,
+ggsave("../manuscript/figures/pareto_demo-1.pdf", pareto_demo,
        width = 5.93, height = 2.33)
 
 # remove lakes below censor threshold ####
@@ -102,7 +102,7 @@ res         <- dplyr::bind_rows(cf_extra, cf)
 predict_area <- ggplot(data = res) +
   geom_line(aes(x = area, y = density, linetype = type)) +
   scale_x_log10() + theme(legend.title = element_blank())
-ggsave("../manuscript/figures/predict_area.pdf", predict_area,
+ggsave("../manuscript/figures/predict_area-1.pdf", predict_area,
        width = 4.28, height = 2.33)
 
 # back-out an estimate of total area
@@ -123,7 +123,7 @@ stack_preds <- data.frame(preds = stack_preds,
 frequentist_uncertainty <- ggplot() +
   geom_line(data = stack_preds, aes(x = area, y = preds, color = type)) +
   scale_x_log10() + xlab("density") + labs(color = "Confidence \n Interval")
-ggsave("../manuscript/figures/frequentist_uncertainty.pdf", frequentist_uncertainty,
+ggsave("../manuscript/figures/frequentist_uncertainty-1.pdf", frequentist_uncertainty,
        width = 5.93, height = 2.33)
 
 # ---- bayesian_model ----
@@ -178,7 +178,7 @@ bayesian_model <- ggplot() + geom_histogram(data = data.frame(alpha = alphas), a
   geom_vline(aes(xintercept = conf_int[c(1,3)]), color = "red") +
   geom_vline(aes(xintercept = conf_int[2])) +
   geom_vline(aes(xintercept = 0.9), linetype = 2)
-ggsave("../manuscript/figures/bayesian_model.pdf", bayesian_model,
+ggsave("../manuscript/figures/bayesian_model-1.pdf", bayesian_model,
        width = 4.43, height = 2.33)
 
 # ---- bayesian_area ----
@@ -207,7 +207,7 @@ bayesian_area <- ggplot() + geom_histogram(data = data.frame(area = area_bayes),
[…]

### 2021-05-18 manuscript/manuscript.tex
correct pdf metadata
typo fix

@@
 pdftitle={Quantifying uncertainty in Pareto lake area estimates},
-pdfsubject={q-bio.NC, q-bio.QM},
+pdfsubject={stat.AP},
 pdfauthor={J.~Stachelek}
@@
 Both studies show that the point at which the distribution mixtures converge can provide valuable inference.
-In the case of lakes, such a convergence point may indicate a change in the data generating process (i.e.
-the point at which lake area are controlled by continent placement rather than fractal landscape morphology \citep{goodchildLakesFractalSurfaces1988, hamiltonEstimationFractalDimension1992}.
+In the case of lakes, such a convergence point may indicate a change in the data generating process such as the point at which lake areas are controlled by continent placement rather than fractal landscape morphology \citep{goodchildLakesFractalSurfaces1988, hamiltonEstimationFractalDimension1992}.
 

### 2021-09-07 manuscript/manuscript.tex
minor grammar fixes

@@
 \author{ \href{https://orcid.org/0000-0002-5924-2464}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}J.~Stachelek}\thanks{\href{https://jsta.rbind.io}{https://jsta.rbind.io}} \\
-	Center for Limnology\\
-	University of Wisconsin-Madison\\
-	Madison, WI 53706 \\
-	\texttt{stachelek@wisc.edu} \\
+	Earth System Observations\\
+	Los Alamos National Laboratory\\
+	Los Alamos, NM, USA, 87544 \\
+	\texttt{jsta@lanl.gov} \\
 }
@@
 \section{Introduction}
-Size is a critical factor determining the contribution of lakes to global biogeochemical cycling and existing evidence suggests that lakes are key components of overall carbon burial, sequestration, and greenhouse gas emissions \citep{delsontroGreenhouseGasEmissions2018, kellerGlobalCarbonBudget2021}.
+Area is a critical factor determining the contribution of lakes to global biogeochemical cycling.
+As such, existing evidence suggests that lakes are key components of overall carbon burial, sequestration, and greenhouse gas emissions \citep{delsontroGreenhouseGasEmissions2018, kellerGlobalCarbonBudget2021}.
 The calculation of such fluxes requires two terms 1) an areal flux rate and 2) the distribution of total lake area.
@@
 In the latter case, uncertainty in total lake area distributions is challenged by the fact that no existing database is a complete census of all lakes \citep{messagerEstimatingVolumeAge2016} and the distribution of individual lake areas spans a range of approximately 7 orders of magnitude.
-The largest lakes are so big they could otherwise be classified as inland seas (> $10^4$ $km^2$) while the smallest are barely larger than a regulation sized soccer field (> $10^{-3}$ $km^2$).
+The largest lakes are so big they could otherwise be classified as inland seas (> $10^4$ $km^2$) while the smallest are barely larger than a regulation sized soccer (football) field (> $10^{-3}$ $km^2$).
 A consequence of these challenges is that uncertainty varies within the distribution as a function of lake size.
@@
 Indeed, many other geomorphological phenomena that are dependent on landform topography such as coastline length are often well-described by fractal generating processes \citep{newman_power_2005}.
-A challenge in modelling such data generating processes for lake areas is large lakes likely follow a different data generating process than that of the smallest lakes.
+A challenge in modelling such data generating processes for lake areas is that large lakes likely follow a different data generating process than that of the smallest lakes.
 Whereas small lakes are constrained by landform topography, large lakes are essentially unconstrained by local landform topography and are instead constrained by the placement and arrangement of continents \citep{goodchildLakesFractalSurfaces1988}.
@@
 They have dealt with the second issue by specifying an ad-hoc cutoff value below which empirically determined lake areas are discarded and subsequently back calculated.
-Hereafter, I refer to this strategy as the cutoff method.
-In the present study, I explore the effects of using the typical frequentist approach for estimating total lake area (i.e.
+Hereafter, I refer to this strategy as the "cutoff" method.
+
+In the present study, I explore two different methods for calculating uncertainty bounds around global lake area.
+First, I calculate these bounds using a frequentist approach for estimating total lake area (i.e.
 the cutoff method) via a simulation study.
-In addition, I demonstrate an approach to produce uncertainty estimates of total lake area in a Bayesian framework.
-I evaluate the effects of the cutoff method on a simulated dataset because it is not sensitive to other potential confounding factors such as heterogeneity of survey effort or data precision.
+Second, I compare these frequentist uncertainty bounds with those calculated using a Bayesian framework.
+I carry out these demonstrations using a simulated dataset so that the ability of each method to recover the "true" parameter values can be evaluated.
+The use of a simulated dataset has the further benefit of avoiding potential confounding factors such as heterogeneity of survey effort or unknown data precision.
 
@@
 Lake areas are typically treated as arising from a scale-invariant fractal generating process \citep{winslowDoesLakeSize2015, downingGlobalAbundanceSize2006, mcdonald_regional_2012, goodchildLakesFractalSurfaces1988, hamiltonEstimationFractalDimension1992}.
-Essentially, this means that the number of lakes in one size class is proportional to the number of lakes in the preceding size class irrespective of their magnitudes.
+This means that the number of lakes in one size class is proportional to the number of lakes in the preceding size class irrespective of their magnitudes.
 The numerical form describing such a process is a power-law function.
@@
 This can be seen from the simulated data in Figure \ref{fig:pareto_demo} where the Pareto pdf contains a lot of noise in the tail but the ccdf appears smoothed.
-The smoothing of the tail is also desirable because it functions as a way of dealing with the truncated nature of lake databases.
+The smoothing of the tail is also desirable because it functions as a way of dealing with the truncated nature of lake databases (i.e.
+the area of large lakes is known exactly).
 The second reason for using the ccdf is that it provides a computational shortcut for estimating the Pareto shape parameter $a$ because it is numerically equivalent to the slope of the ccdf in log-log space \citep{downingGlobalAbundanceSize2006}.
@@
 I simulated a censored lake dataset by excluding lakes smaller than $e^{1}$.
-This excludes (censors) approximately 60\% of the total dataset.
+This excludes (i.e.
+censors) approximately 60\% of the total dataset.
 I approximated the "true" lake area total by constructing the empirical distribution function (edf) of the data which approximates the underlying Pareto cdf \citep{newman_power_2005}.
@@
 
-Instead of the essentially fixed $a$ and total lake area estimates produced by the frequentist approach, I found a substantial variability in both $a$ (95\% CI: 0.86, 0.92) and total area using a Bayesian approach (Fig \ref{fig:bayesian_model}, \ref{fig:bayesian_area}).
+Instead of the essentially fixed $a$ and total lake area estimates produced by the frequentist approach, I found substantial variability in both $a$ (95\% CI: 0.86, 0.92) and total area using a Bayesian approach (Fig \ref{fig:bayesian_model}, \ref{fig:bayesian_area}).
 In particular, the Bayesian 95\% credible intervals for both $a$ and total lake area encapsulate the true values (Fig \ref{fig:bayesian_model}, \ref{fig:bayesian_area}).

### 2021-09-09 scripts/simulation.R
linting

@@ -5,21 +5,21 @@ library(magrittr)
 library(ggplot2)
 library(cowplot)
 
-dpareto <- function(x, a=0.9, b=1) a*b^a/x^(a+1)
-ppareto <- function(x, a=0.9, b=1) (x > b)*(1-(b/x)^a)
-qpareto <- function(u, a=0.9, b=1) b/(1-u)^(1/a)
-rpareto <- function(n, a=0.9, b=1, max = Inf){
+dpareto <- function(x, a = 0.9, b = 1) a * b^a / x^(a + 1)
+ppareto <- function(x, a = 0.9, b = 1) (x > b) * (1 - (b / x)^a)
+qpareto <- function(u, a = 0.9, b = 1) b / (1 - u)^(1 / a)
+rpareto <- function(n, a = 0.9, b = 1, max = Inf) {
   res <- qpareto(runif(n), a, b)
   res <- res[res < max]
 
-  while(length(res) < n){
+  while (length(res) < n) {
     res <- c(res, qpareto(runif(n - length(res)), a, b))
   }
   res
 }
 # (dt <- sample(1:10, 12, replace = TRUE) %>% .[order(.)])
 # (cf <- cumulative_freq(dt))
-cumulative_freq <- function(x){
+cumulative_freq <- function(x) {
   l_n    <- cumsum(rev(table(x)))
   l_area <- rev(as.numeric(names(table(x))))
 
@@ -27,13 +27,13 @@ cumulative_freq <- function(x){
 }
 
 # inv_cumulative_freq(cf)
-inv_cumulative_freq <- function(cf){
-  cf <- cf[order(cf$number),]
+inv_cumulative_freq <- function(cf) {
+  cf <- cf[order(cf$number), ]
   rep(cf$area, c(cf$number[1], diff(cf$number)))
 }
 
 # simulate random pareto draws
-if(!file.exists("../data/y.rds")){
+if (!file.exists("../data/y.rds")) {
   y <- rpareto(10000, max = 81935.7) # cap at the area of Lake Superior
   saveRDS(y, "../data/y.rds")
 }
@@ -43,8 +43,10 @@ y <- readRDS("../data/y.rds")
 pareto_demo <- plot_grid(
   # individual binning
   hist(log(y), plot = FALSE, n = 100) %>%
-  {data.frame(x = .$breaks[-1], samples = log(.$counts))} %>%
-    ggplot + geom_line(aes(x, samples)) +
+    {
[…]

### 2021-09-09 manuscript/manuscript.tex
more precise lower area estimate

@@
 It is likely that such underestimation, would increase with a greater proportion of censoring.
-This may explain the steady increase in estimates of global lake area through time from approximately 3 to 5 million $km^{2}$ as lake area databases have improved their accuracy \citep{lehnerDevelopmentValidationGlobal2004, downingGlobalAbundanceSize2006, verpoorterGlobalInventoryLakes2014}.
+This may explain the steady increase in estimates of global lake area through time from approximately 2.5 to 5 million $km^{2}$ as lake area databases have improved their accuracy \citep{lehnerDevelopmentValidationGlobal2004, downingGlobalAbundanceSize2006, verpoorterGlobalInventoryLakes2014}.
 Future work on estimating global lake area should consider implementing a sensitivity analysis looking at the response of total area estimates to variation in the degree of censoring.

### 2021-09-09 scripts/simulation.R
file path fix

@@ -33,12 +33,12 @@ inv_cumulative_freq <- function(cf) {
 }
 
 # simulate random pareto draws
-if (!file.exists("../data/y.rds")) {
+if (!file.exists("data/y.rds")) {
   y <- rpareto(10000, max = 81935.7) # cap at the area of Lake Superior
-  saveRDS(y, "../data/y.rds")
+  saveRDS(y, "data/y.rds")
 }
 #
-y <- readRDS("../data/y.rds")
+y <- readRDS("data/y.rds")
 
 pareto_demo <- plot_grid(
   # individual binning
@@ -59,7 +59,7 @@ pareto_demo <- plot_grid(
     theme_minimal() + scale_x_log10() + scale_y_log10() +
     ylab("samples with value > x") + xlab("x"),
   nrow = 1)
-ggsave("../manuscript/figures/pareto_demo-1.pdf", pareto_demo,
+ggsave("manuscript/figures/pareto_demo-1.pdf", pareto_demo,
   width = 5.93, height = 2.33)
 
 # remove lakes below censor threshold ####
@@ -85,7 +85,7 @@ total_empirical <- sum(inv_cumulative_freq(cumulative_freq(y)))
 cf         <- cumulative_freq(y_censored)
 cf$density <- log(cf$number) - max(log(cf$number))
 cf$type    <- "empirical"
-fit_lm        <- lm(density ~ log(area), data = cf)
+fit_lm     <- lm(density ~ log(area), data = cf)
 
 # extrapolate censored lake counts
 cf_extra <- data.frame(area = exp(seq(
@@ -106,7 +106,7 @@ res         <- dplyr::bind_rows(cf_extra, cf)
 predict_area <- ggplot(data = res) +
   geom_line(aes(x = area, y = density, linetype = type)) +
   scale_x_log10() + theme(legend.title = element_blank())
-ggsave("../manuscript/figures/predict_area-1.pdf", predict_area,
+ggsave("manuscript/figures/predict_area-1.pdf", predict_area,
   width = 4.28, height = 2.33)
 
 # back-out an estimate of total area
@@ -127,7 +127,7 @@ stack_preds <- data.frame(preds = stack_preds,
 frequentist_uncertainty <- ggplot() +
   geom_line(data = stack_preds, aes(x = area, y = preds, color = type)) +
   scale_x_log10() + xlab("density") + labs(color = "Confidence \n Interval")
[…]

### 2021-09-09 scripts/simulation.R
linting

@@ -127,8 +127,8 @@ stack_preds <- data.frame(preds = stack_preds,
 frequentist_uncertainty <- ggplot() +
   geom_line(data = stack_preds, aes(x = area, y = preds, color = type)) +
   scale_x_log10() + xlab("density") + labs(color = "Confidence \n Interval")
-ggsave("manuscript/figures/frequentist_uncertainty-1.pdf", frequentist_uncertainty,
-  width = 5.93, height = 2.33)
+ggsave("manuscript/figures/frequentist_uncertainty-1.pdf",
+  frequentist_uncertainty, width = 5.93, height = 2.33)
 
 # ---- bayesian_model ----
 # https://github.com/stan-dev/example-models/blob/master/bugs_examples/vol3/fire/fire.stan
@@ -178,7 +178,8 @@ if (!file.exists("data/alphas.rds")) {
 alphas <- readRDS("data/alphas.rds")
 
 conf_int <- quantile(alphas, probs = c(0.025, 0.5, .975))
-bayesian_model <- ggplot() + geom_histogram(data = data.frame(alpha = alphas), aes(x = alpha)) +
+bayesian_model <- ggplot() +
+  geom_histogram(data = data.frame(alpha = alphas), aes(x = alpha)) +
   geom_vline(aes(xintercept = conf_int[c(1, 3)]), color = "red") +
   geom_vline(aes(xintercept = conf_int[2])) +
   geom_vline(aes(xintercept = 0.9), linetype = 2)
@@ -192,7 +193,8 @@ if (!file.exists("data/area_bayes.rds")) {
   # back-out an estimate of total area
   area_bayes <- sapply(alphas, function(a) {
     cf_extra_bayes <- select(cf_extra, area)
-    cf_extra_bayes$density <- (log(cf_extra_bayes$area) * (a * -1)) + log(min(y_censored)) - 0.1
+    cf_extra_bayes$density <- (log(cf_extra_bayes$area) * (a * -1)) +
+      log(min(y_censored)) - 0.1
     cf_extra_bayes$type    <- "predicted_bayes"
     cf_extra_bayes$number  <- exp(cf_extra_bayes$density + max(log(cf$number)))
     res <- dplyr::bind_rows(cf_extra_bayes, cf)
@@ -207,7 +209,8 @@ if (!file.exists("data/area_bayes.rds")) {
 area_bayes <- readRDS("data/area_bayes.rds")
 
[…]

### 2021-09-09 scripts/simulation.R
read-in hydrolakes data, increase simulation to match hydrolakes reported total

@@ -4,6 +4,8 @@ library(dplyr)
 library(magrittr)
 library(ggplot2)
 library(cowplot)
+library(rstan)
+library(sf)
 
 dpareto <- function(x, a = 0.9, b = 1) a * b^a / x^(a + 1)
 ppareto <- function(x, a = 0.9, b = 1) (x > b) * (1 - (b / x)^a)
@@ -32,9 +34,52 @@ inv_cumulative_freq <- function(cf) {
   rep(cf$area, c(cf$number[1], diff(cf$number)))
 }
 
+
+#' Read a config.py file in R
+#'
+#' @param file a file path
+#' @export
+#' @examples \dontrun{
+#' # See jsta/rjsta @ Github
+#' dir(config()$a)
+#' }
+config <- function(file = "config.py") {
+  res <- suppressWarnings(readLines(file))
+
+  # remove non equal sign lines
+  is_assignment <- as.logical(sapply(res, function(x) grep("=", x) >= 1))
+  is_assignment <- sapply(is_assignment,
+    function(x) ifelse(is.na(x), FALSE, TRUE))
+  res <- res[is_assignment]
+
+  res <- strsplit(res, " ")
+  keys <- unlist(lapply(res, function(x) x[[1]][[1]]))
+  values <- unlist(lapply(res, function(x) x[[3]][[1]]))
+  # keys <- unlist(lapply(res, function(x) strsplit(x, "=")[[1]][[1]]))
+  # values <- unlist(lapply(res, function(x) strsplit(x, "=")[[1]][[2]]))
+
+  res <- setNames(data.frame(t(values)), keys)
+
+  norm_file_path <- function(x) {
+    gsub('\\"', "", gsub("\\\\\\\\", "\\\\", x))
+  }
+
+  data.frame(t(apply(res, 2, norm_file_path)))
+}
+
+# read-in hydrolakes data
+data_path_hydrolakes <- config()$data_path_hydrolakes
+data_path_hydrolakes <- file.path(data_path_hydrolakes,
+  "HydroLAKES_polys_v10_shp", "HydroLAKES_polys_v10.shp")
+area_hydrolakes <- st_read(data_path_hydrolakes,
+  query = "SELECT Lake_area FROM \"HydroLAKES_polys_v10\"") %>%
+  st_drop_geometry()
+# y <- area_hydrolakes$Lake_area
+
 # simulate random pareto draws
+# unlink("data/y.rds")
 if (!file.exists("data/y.rds")) {
[…]

### 2021-09-12 scripts/simulation.R
move simulation closer to hydrolakes total

@@ -79,7 +79,7 @@ area_hydrolakes <- st_read(data_path_hydrolakes,
 # simulate random pareto draws
 # unlink("data/y.rds")
 if (!file.exists("data/y.rds")) {
-  y <- rpareto(140000, max = 81935.7) # cap at the area of Lake Superior
+  y <- rpareto(153000, max = 81935.7) # cap at the area of Lake Superior
   saveRDS(y, "data/y.rds")
 }
 #

### 2021-09-12 scripts/simulation.R
fix file paths, silence startup text

@@ -1,11 +1,15 @@
 # ---- pareto_demo ----
 # setwd("scripts")
-library(dplyr)
+suppressMessages(library(dplyr))
 library(magrittr)
 library(ggplot2)
 library(cowplot)
-library(rstan)
-library(sf)
+suppressMessages(library(rstan))
+suppressMessages(library(sf))
+# setwd("scripts")
+
+options(mc.cores = parallel::detectCores())
+rstan_options(auto_write = TRUE)
 
 dpareto <- function(x, a = 0.9, b = 1) a * b^a / x^(a + 1)
 ppareto <- function(x, a = 0.9, b = 1) (x > b) * (1 - (b / x)^a)
@@ -68,43 +72,53 @@ config <- function(file = "config.py") {
 }
 
 # read-in hydrolakes data
-data_path_hydrolakes <- config()$data_path_hydrolakes
+data_path_hydrolakes <- config("../config.py")$data_path_hydrolakes
 data_path_hydrolakes <- file.path(data_path_hydrolakes,
   "HydroLAKES_polys_v10_shp", "HydroLAKES_polys_v10.shp")
 area_hydrolakes <- st_read(data_path_hydrolakes,
-  query = "SELECT Lake_area FROM \"HydroLAKES_polys_v10\"") %>%
+  query = "SELECT Lake_area FROM \"HydroLAKES_polys_v10\"", quiet = TRUE) %>%
   st_drop_geometry()
-# y <- area_hydrolakes$Lake_area
+area_hydrolakes <- area_hydrolakes$Lake_area
 
 # simulate random pareto draws
-# unlink("data/y.rds")
-if (!file.exists("data/y.rds")) {
+# unlink("../data/y.rds")
+if (!file.exists("../data/y.rds")) {
   y <- rpareto(153000, max = 81935.7) # cap at the area of Lake Superior
-  saveRDS(y, "data/y.rds")
+  saveRDS(y, "../data/y.rds")
 }
 #
-y <- readRDS("data/y.rds")
+y <- readRDS("../data/y.rds")
 
-pareto_demo <- plot_grid(
-  # individual binning
-  hist(log(y), plot = FALSE, n = 100) %>%
-    {
-      data.frame(x = .$breaks[-1], samples = log(.$counts))
-    } %>%
-    ggplot() + geom_line(aes(x, samples)) +
-    theme_minimal(),
-  # log binning
-  # hist(y, plot = FALSE, breaks = 10^(seq(from = 0, to = 6, by = 0.6))) %>%
[…]

### 2021-09-12 scripts/simulation.R
better variable naming

@@ -128,13 +128,14 @@ y_censored <- y[log(y) > 1]
 # ---- recover_area ----
 
 # back out total "area" from individual binning
-test <- hist(log(y), plot = FALSE, n = 10000) %>%
+total_area <- hist(log(y), plot = FALSE, n = 10000) %>%
   {
     data.frame(x = .$breaks[-1], samples = log(.$counts))
   } %>%
   filter(samples > -Inf)
 
-sum(exp(test$x) * exp(test$samples))
+total_area <- sum(exp(total_area$x) * exp(total_area$samples))
+# print(total_area)
 
 # ---- predict_area ----
 # back out total "area" from cumulative frequency

### 2021-09-12 scripts/simulation.R
rework plot 1 to compare simulation data with hydrolakes

@@ -8,6 +8,7 @@ suppressMessages(library(rstan))
 suppressMessages(library(sf))
 # setwd("scripts")
 
+set.seed(234)
 options(mc.cores = parallel::detectCores())
 rstan_options(auto_write = TRUE)
 
@@ -89,34 +90,43 @@ if (!file.exists("../data/y.rds")) {
 #
 y <- readRDS("../data/y.rds")
 
-pareto_plot <- function(dt) {
-  plot_grid(
-    # individual binning
-    hist(log(dt), plot = FALSE, n = 100) %>%
-      {
-        data.frame(x = .$breaks[-1], samples = log(.$counts))
-      } %>%
-      ggplot() + geom_line(aes(x, samples)) +
-      theme_minimal(),
-    # log binning
-    # hist(y, plot = FALSE, breaks = 10^(seq(from = 0, to = 6, by = 0.6))) %>%
-    # {data.frame(x = .$breaks[-1], samples = .$counts + 0.01)} %>%
-    #   ggplot + geom_line(aes(x, samples)) +
-    #   theme_minimal() + scale_x_log10() + scale_y_log10(),
-    # cumulative binning
-    cumulative_freq(dt) %>%
-      ggplot() + geom_line(aes(area, number)) +
-      theme_minimal() + scale_x_log10() + scale_y_log10() +
-      ylab("samples with value > x") + xlab("x"),
-    nrow = 1)
+pareto_plot_prep <- function(x) {
+  individual_binning <- hist(log(x), plot = FALSE, n = 100) %>%  {
+    data.frame(x = .$breaks[-1], samples = log(.$counts))
+  }
+  cumulative_binning <- cumulative_freq(x)
+
+  list(individual_binning = individual_binning,
+    cumulative_binning = cumulative_binning)
 }
 
-# plot_grid(
-#   pareto_plot(y),
-#   pareto_plot(area_hydrolakes),
-#   nrow = 2, ncol = 1
-# )
-pareto_demo <- pareto_plot(y)
+pareto_plot <- function(individual_binning, cumulative_binning) {
+  gg_individual <- ggplot(data = individual_binning) +
+    geom_line(aes(x, samples, color = name)) +
+    theme_minimal() +
+    theme(legend.position = "none")
+
+  gg_cumulative <- ggplot(data = cumulative_binning) +
+    geom_line(aes(area, number, color = name)) +
[…]

### 2021-09-13 scripts/simulation.R
increase model iterations, set figure xlims

@@ -6,6 +6,7 @@ library(ggplot2)
 library(cowplot)
 suppressMessages(library(rstan))
 suppressMessages(library(sf))
+library(tidybayes)
 # setwd("scripts")
 
 set.seed(234)
@@ -228,10 +229,12 @@ model {
 }
 "
 
+# unlink("../data/pareto_bayes.rds"); unlink("../data/alphas.rds"); unlink("../data/area_bayes.rds")
+
 if (!file.exists("../data/pareto_bayes.rds")) {
   fit <- stan(model_code = pareto_model,
     data = list(N = length(y_censored), x = y_censored),
-    iter = 8000)
+    iter = 15000)
 
   # print(fit)
   # plot(fit, pars = "alpha")
@@ -240,7 +243,7 @@ if (!file.exists("../data/pareto_bayes.rds")) {
 fit <- readRDS("../data/pareto_bayes.rds")
 
 if (!file.exists("../data/alphas.rds")) {
-  alphas <- tidybayes::spread_samples(fit, alpha)$alpha
+  alphas <- tidybayes::spread_draws(fit, alpha)$alpha
   saveRDS(alphas, "../data/alphas.rds")
 }
 alphas <- readRDS("../data/alphas.rds")
@@ -250,7 +253,10 @@ bayesian_model <- ggplot() +
   geom_histogram(data = data.frame(alpha = alphas), aes(x = alpha)) +
   geom_vline(aes(xintercept = conf_int[c(1, 3)]), color = "red") +
   geom_vline(aes(xintercept = conf_int[2])) +
-  geom_vline(aes(xintercept = 0.9), linetype = 2)
+  geom_vline(aes(xintercept = 0.9), linetype = 2) +
+  xlim(
+    conf_int[2] - ((conf_int[2] - conf_int[1]) * 3),
+    conf_int[2] + ((conf_int[3] - conf_int[2]) * 3))
 ggsave("../manuscript/figures/bayesian_model-1.pdf", bayesian_model,
   width = 4.43, height = 2.33)
 
@@ -281,7 +287,9 @@ bayesian_area <- ggplot() +
   geom_histogram(data = data.frame(area = area_bayes), aes(x = area)) +
   geom_vline(aes(xintercept = conf_int[c(1, 3)]), color = "red") +
   geom_vline(aes(xintercept = conf_int[2])) +
-  geom_vline(aes(xintercept = total_empirical), linetype = 2)
[…]

### 2021-09-13 manuscript/manuscript.tex
edits
rm personal website

--- 2021-09-07 manuscript/manuscript.tex
@@
 
-\author{ \href{https://orcid.org/0000-0002-5924-2464}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}J.~Stachelek}\thanks{\href{https://jsta.rbind.io}{https://jsta.rbind.io}} \\
+\author{ \href{https://orcid.org/0000-0002-5924-2464}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}J.~Stachelek} \\
 	Earth System Observations\\

### 2021-09-13 scripts/simulation.R
use seq_len

@@ -184,11 +184,11 @@ ggsave("../manuscript/figures/predict_area-1.pdf", predict_area,
 total_predicted <- sum(inv_cumulative_freq(res))
 
 # ---- frequentist_uncertainty ----
-stack_preds <- c(res$density[1:nrow(cf_extra)],
-  res$lower[1:nrow(cf_extra)],
-  res$upper[1:nrow(cf_extra)])
+stack_preds <- c(res$density[seq_len(nrow(cf_extra))],
+  res$lower[seq_len(nrow(cf_extra))],
+  res$upper[seq_len(nrow(cf_extra))])
 stack_preds <- data.frame(preds = stack_preds,
-  area  = rep(res$area[1:nrow(cf_extra)], times = 3),
+  area  = rep(res$area[seq_len(nrow(cf_extra))], times = 3),
   type  = rep(c("50", "2.5", "97.5"),
     each = nrow(cf_extra)))
 # res$number_lower[(nrow(cf_extra) + 1):nrow(res)] <-

### 2021-09-13 scripts/simulation.R
more intutive axis labels for fig 1

@@ -92,6 +92,7 @@ if (!file.exists("../data/y.rds")) {
 y <- readRDS("../data/y.rds")
 
 pareto_plot_prep <- function(x) {
+  # x <- y
   individual_binning <- hist(log(x), plot = FALSE, n = 100) %>%  {
     data.frame(x = .$breaks[-1], samples = log(.$counts))
   }
@@ -104,13 +105,14 @@ pareto_plot_prep <- function(x) {
 pareto_plot <- function(individual_binning, cumulative_binning) {
   gg_individual <- ggplot(data = individual_binning) +
     geom_line(aes(x, samples, color = name)) +
+    ylab("log(n)") + xlab("log(area)") +
     theme_minimal() +
     theme(legend.position = "none")
 
   gg_cumulative <- ggplot(data = cumulative_binning) +
     geom_line(aes(area, number, color = name)) +
     theme_minimal() + scale_x_log10() + scale_y_log10() +
-    ylab("samples with value > x") + xlab("x") +
+    ylab("n > area") + xlab("area") +
     labs(color = "")
 
   plot_grid(gg_individual, gg_cumulative,

### 2021-09-28 scripts/01_censoring.R
censored predictions match hydrolakes finally

@@ -1,17 +1,43 @@
 # setwd("scripts")
-source("utils.R")
+source("scripts/utils.R")
 
-y <- readRDS("../data/y.rds")
+y <- readRDS("data/y.rds")
+y_raw <- y$y_raw
+# hist(log(y_raw))
 
-# remove lakes below censor threshold ####
-# hist(log(y))
-y_censored <- y[log(y) > 1]
-saveRDS(y_censored, "../data/y_censored.rds")
+# remove lakes below censor threshold ----
+y_censored_raw <- y_raw[log(y_raw) > 1]
+y_censored_cumulative <- mutate(
+  cumulative_freq(y_censored_raw), name = "censored")
+y_censored_individual <- individual_freq(y_censored_raw)
 
-# recover_area ----
+#  compare hydrolakes and y_censored_cumulative
+hydrolakes <- readRDS("data/hydrolakes.rds")
+rbind(y_censored_cumulative, hydrolakes$hydrolakes_cumulative) %>%
+  ggplot() +
+  geom_line(aes(area, number, color = name)) +
+  theme_minimal() + scale_x_log10() + scale_y_log10() +
+  ylab("n > area") + xlab("area") +
+  labs(color = "")
 
-# back out total "area" from individual binning
-total_area <- hist(log(y), plot = FALSE, n = 10000) %>%
+# tail(y_censored_cumulative)
+# tail(y$y_cumulative)
+# dplyr::filter(y$y_cumulative, area > 2.718311) %>% tail()
+
+# range(y_censored_raw)
+# hist(log(y_raw))
+# hist(log(y_censored_raw))
+
+#
+
+# y <- c(y,
+#   list(
+#     y_censored_raw = y_censored_raw,
+#     y_censored_cumulative = y_censored_cumulative))
+# saveRDS(y, "data/y.rds")
+
+# back out total "area" from individual binning ----
+total_area <- hist(log(y_raw), plot = FALSE, n = 10000) %>%
   {
     data.frame(x = .$breaks[-1], samples = log(.$counts))
   } %>%
@@ -20,12 +46,12 @@ total_area <- hist(log(y), plot = FALSE, n = 10000) %>%
 total_area <- sum(exp(total_area$x) * exp(total_area$samples))
 # print(total_area)
 
-# predict_area ----
-# back out total "area" from cumulative frequency
-total_empirical <- sum(inv_cumulative_freq(cumulative_freq(y)))
+# back out total "area" from cumulative frequency ----
[…]

### 2021-09-28 scripts/02_frequentist.R
file path fix

@@ -1,8 +1,7 @@
-# setwd("scripts")
-source("utils.R")
+source("scripts/utils.R")
 
-res <- readRDS("../data/res.rds")
-cf_extra <- readRDS("../data/cf_extra.rds")
+res <- readRDS("data/res.rds")
+cf_extra <- readRDS("data/cf_extra.rds")
 
 # ---- frequentist_uncertainty ----
 stack_preds <- c(res$density[seq_len(nrow(cf_extra))],
@@ -19,5 +18,6 @@ stack_preds <- data.frame(preds = stack_preds,
 frequentist_uncertainty <- ggplot() +
   geom_line(data = stack_preds, aes(x = area, y = preds, color = type)) +
   scale_x_log10() + xlab("density") + labs(color = "Confidence \n Interval")
-ggsave("../manuscript/figures/frequentist_uncertainty-1.pdf",
+
+ggsave("manuscript/figures/frequentist_uncertainty-1.pdf",
   frequentist_uncertainty, width = 5.93, height = 2.33)

### 2021-10-06 scripts/utils.R
load rstan

@@ -7,6 +7,7 @@ library(cowplot)
 suppressMessages(library(rstan))
 suppressMessages(library(sf))
 library(tidybayes)
+library(rstan)
 
 # functions ----
 

### 2021-10-06 scripts/01_censoring.R
save cf object

@@ -80,6 +80,7 @@ res <- dplyr::bind_rows(cf_extra, cf) %>%
 
 saveRDS(res, "data/res.rds")
 saveRDS(cf_extra, "data/cf_extra.rds")
+saveRDS(cf, "data/cf.rds")
 
 # compare hydrolakes and censored predictions
 hydrolakes <- readRDS("data/hydrolakes.rds")

### 2021-10-06 scripts/03_bayesian.R
fix paths

@@ -1,10 +1,10 @@
-# setwd("scripts")
-source("utils.R")
+source("scripts/utils.R")
 options(mc.cores = parallel::detectCores())
 rstan_options(auto_write = TRUE)
 
-y_censored <- readRDS("../data/y_censored.rds")
-cf_extra <- readRDS("../data/cf_extra.rds")
+y_censored <- readRDS("data/y_censored.rds")
+cf_extra <- readRDS("data/cf_extra.rds")
+cf <- readRDS("data/cf.rds")
 
 # ---- bayesian_model ----
 # nolint start
@@ -40,22 +40,22 @@ model {
 # unlink("../data/pareto_bayes.rds"); unlink("../data/alphas.rds"); unlink("../data/area_bayes.rds")
 # nolint end
 
-if (!file.exists("../data/pareto_bayes.rds")) {
+if (!file.exists("data/pareto_bayes.rds")) {
   fit <- stan(model_code = pareto_model,
     data = list(N = length(y_censored), x = y_censored),
     iter = 15000)
 
   # print(fit)
   # plot(fit, pars = "alpha")
-  saveRDS(fit, "../data/pareto_bayes.rds")
+  saveRDS(fit, "data/pareto_bayes.rds")
 }
-fit <- readRDS("../data/pareto_bayes.rds")
+fit <- readRDS("data/pareto_bayes.rds")
 
-if (!file.exists("../data/alphas.rds")) {
+if (!file.exists("data/alphas.rds")) {
   alphas <- tidybayes::spread_draws(fit, alpha)$alpha
-  saveRDS(alphas, "../data/alphas.rds")
+  saveRDS(alphas, "data/alphas.rds")
 }
-alphas <- readRDS("../data/alphas.rds")
+alphas <- readRDS("data/alphas.rds")
 
 conf_int <- quantile(alphas, probs = c(0.025, 0.5, .975))
 bayesian_model <- ggplot() +
@@ -66,12 +66,19 @@ bayesian_model <- ggplot() +
   xlim(
     conf_int[2] - ((conf_int[2] - conf_int[1]) * 3),
     conf_int[2] + ((conf_int[3] - conf_int[2]) * 3))
-ggsave("../manuscript/figures/bayesian_model-1.pdf", bayesian_model,
+ggsave("manuscript/figures/bayesian_model-1.pdf", bayesian_model,
   width = 4.43, height = 2.33)
 
 # ---- bayesian_area ----
 
-if (!file.exists("../data/area_bayes.rds")) {
[…]

### 2021-10-06 scripts/01_censoring.R
manage data for bayesian fitting

@@ -10,6 +10,7 @@ y_censored_raw <- y_raw[log(y_raw) > 1]
 y_censored_cumulative <- mutate(
   cumulative_freq(y_censored_raw), name = "censored")
 y_censored_individual <- individual_freq(y_censored_raw)
+saveRDS(y_censored_raw, "data/y_censored.rds")
 
 #  compare hydrolakes and y_censored_cumulative
 hydrolakes <- readRDS("data/hydrolakes.rds")

### 2021-10-06 scripts/03_bayesian.R

@@ -82,6 +82,7 @@ if (!file.exists("data/area_bayes.rds")) {
   # find estimated density of censored lakes given alpha
   # back-out an estimate of total area
   area_bayes <- sapply(alphas, function(a) {
+    # a <- alphas[1]
     cf_extra_bayes <- select(cf_extra, area)
     cf_extra_bayes$density <- (log(cf_extra_bayes$area) * (a * -1)) +
       log(min(y_censored)) - 0.1

### 2021-10-06 scripts/01_censoring.R
linting

@@ -66,7 +66,9 @@ fit_lm     <- lm(density ~ log(area), data = cf)
 cf_extra <- data.frame(area = exp(seq(
   from = min(log(y_raw)),
   to   = min(log(y_censored_raw)),
-  by   = abs(mean(diff(log(cf$area)))))))
+  by   = abs(mean(diff(log(cf$area))))
+  ))
+  )
 
 preds <- data.frame(predict(fit_lm, cf_extra, interval = "confidence"))
 preds <- setNames(preds, c("density", "lower", "upper"))

### 2021-10-06 scripts/03_bayesian.R
better xlim

@@ -105,7 +105,9 @@ bayesian_area <- ggplot() +
   geom_vline(aes(xintercept = conf_int[c(1, 3)]), color = "red") +
   geom_vline(aes(xintercept = conf_int[2])) +
   geom_vline(aes(xintercept = total_empirical), linetype = 2) +
-  xlim(conf_int[2] - ((conf_int[2] - conf_int[1]) * 2), total_empirical
+  xlim(
+  conf_int[2] - ((conf_int[2] - conf_int[1]) * 2),
+  conf_int[3] + ((conf_int[3] - conf_int[2]) * 2)
   )
 ggsave("manuscript/figures/bayesian_area-1.pdf", bayesian_area,
   width = 4.36, height = 2.33)

### 2021-10-06 manuscript/manuscript.tex
censored predictions match hydrolakes finally
update title
cleanup
frequentist table placeholder

@@
 \usepackage{microtype}      % microtypography
-\usepackage{lipsum}		% Can be removed after putting your text content
 \usepackage{graphicx}
@@
 
-\title{Quantifying uncertainty in Pareto lake area estimates}
+\title{Quantifying uncertainty in Pareto estimates of global lake area}
 
 %\date{September 9, 1985}	% Here you can change the date presented in the paper title
-%\date{} 					% Or removing it
+\date{} 					% Or removing it
 
@@
 In addition, estimates of total lake area are somewhat close to the "true" value (Table \ref{tab:pred_table}).
-However, uncertainty around the frequentist estimates is unrealistically small (Figure \ref{fig:frequentist_uncertainty}).
+However, uncertainty around the frequentist estimates is unrealistically small (Figure \ref{tab:frequentist_uncertainty}).
 
@@
 	\centering
-	\includegraphics{figures/predict_area-1}
+	\includegraphics{figures/predict_censor-1}
 	\caption{Censored lake area edf (solid line) and cdf estimate (dashed line).}
-	\label{fig:predict_area}
+	\label{fig:predict_censor}
 \end{figure}
 
-\begin{figure}
+% \begin{figure}
+% 	\centering
+% 	\includegraphics{figures/frequentist_uncertainty-1}
+% 	\caption{Confidence interval of frequentist density predictions for censored data (the dashed portion in Figure 2. The confidence interval here has essentially no width.}
+% 	\label{tab:frequentist_uncertainty}
+% \end{figure}
+\begin{table}
+	\caption{Frequentist uncertainty}
 	\centering
-	\includegraphics{figures/frequentist_uncertainty-1}
-	\caption{Confidence interval of frequentist density predictions for censored data (the dashed portion in Figure 2. The confidence interval here has essentially no width.}
-	\label{fig:frequentist_uncertainty}
-\end{figure}
+	\begin{tabular}{ccc}
+		\toprule
+		Q5 ($km^2$) & Q50 ($km^2$) & Q95 ($km^2$) \\
+		\midrule
+		2.271e+05          & 2.263e+05   & 123     \\
+		\bottomrule
+	\end{tabular}
+	\label{tab:frequentist_uncertainty}
+\end{table}
 
@@
 
-I have shown that the typical frequentist cutoff method produces reasonable estimates of the density of small censored lakes but that it does not capture uncertainty in total lake area (Table \ref{tab:pred_table}, Figure \ref{fig:frequentist_uncertainty}).
+I have shown that the typical frequentist cutoff method produces reasonable estimates of the density of small censored lakes but that it does not capture uncertainty in total lake area (Table \ref{tab:pred_table}, Figure \ref{tab:frequentist_uncertainty}).
 Furthermore, I have shown that models fit using a Bayesian approach indicate substantial uncertainty in both total lake area and the underlying Pareto shape parameter $a$ used to derive these estimates (Figure \ref{fig:bayesian_model}, \ref{fig:bayesian_area}).

### 2021-10-20 manuscript/manuscript.tex
minor text edits

@@
 Because we do not have a complete census of all lakes, upscaling estimates of such processes at broad spatial scales requires the use of lake size-abundance distributions rather than empirical measurements of area.
-Existing lake census efforts are incomplete such that as lakes become smaller they are more likely to by omitted either because they are too small to be resolved from remote sensing products or because of limited ground surveying effort (i.e.
+Existing lake census efforts are incomplete such that as lakes become smaller they are more likely to be omitted either because they are too small to be resolved from remote sensing products or because of limited ground surveying effort (i.e.
 "censoring" of small lakes relative to large lakes).
-
 The present study explores one potential shortcoming of prior approaches estimating global lake area using lake size-abundance distributions.

### 2021-10-20 manuscript/figures.Rmd
towards submission

@@
 
-```{r 01_iris, fig.cap="Sepal length versus width", echo=FALSE, fig.width=1,out.height="4in"}
-knitr::include_graphics("../figures/01_iris-1.pdf")
+```{r fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution
+function.", echo=FALSE, fig.width=1,out.height="4in"}
+knitr::include_graphics("manuscript/figures/pareto_demo-1.pdf")
 ```

### 2021-10-20 manuscript/tables/reprex.R

@@ -1,0 +1,32 @@
+library(reprex)
+
+tmp <- file.path(tempdir(), "foofy")
+dir.create(tmp)
+reprex({
+    #+ setup, include = FALSE
+    knitr::opts_chunk$set(results=FALSE)  
+"
+data {
+  int<lower=0> N;
+  real x[N];
+}
+parameters {
+  real<lower=0> alpha;
+  real<lower=0> theta;
+}
+model {
+  real lpa[N];
+
+  theta ~ gamma(1, 3);
+  alpha ~ gamma(1, 3);
+
+  for (i in 1:N) {
+    lpa[i] = pareto_lpdf(x[i] | theta, alpha);
+  }
+
+  target += sum(lpa);
+}
+"
+  
+  }, wd = ".", advertise = FALSE)
+list.files(dirname(tmp), pattern = "foofy")

### 2021-10-20 scripts/03_bayesian.R
use realistic inits

@@ -42,9 +42,8 @@ model {
 
 
 if (!file.exists("data/pareto_bayes.rds")) {
-  alpha_inits <- 0.9
   init_f <- function() {
-    list(alpha = alpha_inits);
+    list(alpha = runif(1, 0.66, 1.334)) # from Downing et al. 2006
   }
 
   fit <- stan(model_code = pareto_model,
@@ -114,8 +113,8 @@ bayesian_area <- ggplot() +
   geom_vline(aes(xintercept = conf_int[2])) +
   geom_vline(aes(xintercept = total_empirical), linetype = 2) +
   xlim(
-  conf_int[2] - ((conf_int[2] - conf_int[1]) * 2),
-  conf_int[3] + ((conf_int[3] - conf_int[2]) * 2)
+    conf_int[2] - ((conf_int[2] - conf_int[1]) * 2),
+    conf_int[3] + ((conf_int[3] - conf_int[2]) * 2)
   )
 ggsave("manuscript/figures/bayesian_area-1.pdf", bayesian_area,
   width = 4.36, height = 2.33)

### 2021-10-20 manuscript/figures.Rmd
read stan model from file

--- 2018-10-05 manuscript/figures.Rmd
@@
 
-```{r 01_iris, fig.cap="Sepal length versus width", echo=FALSE, fig.width=1,out.height="4in"}
-knitr::include_graphics("../figures/01_iris-1.pdf")
+\pagebreak
+
+```{r fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.", echo=FALSE, fig.width=1,out.height="4in"}
+knitr::include_graphics("../manuscript/figures/pareto_demo-1.pdf")
 ```
+
+```{r fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.", echo=FALSE, fig.width=1,out.height="4in"}
+knitr::include_graphics("../manuscript/figures/pareto_demo-1.pdf")
+```
+
+```{r fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.", echo=FALSE, fig.width=1,out.height="4in"}
+knitr::include_graphics("../manuscript/figures/pareto_demo-1.pdf")
+```
+
+```{r fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.", echo=FALSE, fig.width=1,out.height="4in"}
+knitr::include_graphics("../manuscript/figures/pareto_demo-1.pdf")
+```
+
+```{r fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.", echo=FALSE, fig.width=1,out.height="4in"}
+knitr::include_graphics("../manuscript/figures/pareto_demo-1.pdf")
+```

### 2021-10-20 scripts/03_bayesian.R

@@ -46,7 +46,7 @@ if (!file.exists("data/pareto_bayes.rds")) {
     list(alpha = runif(1, 0.66, 1.334)) # from Downing et al. 2006
   }
 
-  fit <- stan(model_code = pareto_model,
+  fit <- stan(file = "pareto_model.stan",
     data = list(N = length(y_censored), x = y_censored),
     iter = 25000,
     control = list(adapt_delta = 0.9),

### 2021-10-21 manuscript/tables/table_1.R
fill out table1

@@ -1,0 +1,31 @@
+source("scripts/utils.R")
+library(flextable)
+
+y_raw <- readRDS("data/y.rds")$y_raw
+total_empirical <- sum(inv_cumulative_freq(cumulative_freq(y_raw)))
+
+res <- readRDS("data/res.rds")
+total_predicted <- data.frame(
+  q50 = sum(inv_cumulative_freq(res)), 
+  q5 = sum(inv_cumulative_freq(
+    mutate(mutate(res, number_lower = coalesce(number_lower, number)), number = number_lower)
+    )),
+  q95 = sum(inv_cumulative_freq(
+    mutate(mutate(res, number_upper = coalesce(number_upper, number)), number = number_upper)
+    )) 
+  )
+
+area_bayes <- readRDS("data/area_bayes.rds")
+conf_int <- quantile(area_bayes, probs = c(0.025, 0.5, .975))
+
+res <- tibble::tribble(
+  ~ name,        ~ Q50, ~ Q5, ~ Q95,
+    "True",        total_empirical, NA, NA, 
+    "Frequentist", total_predicted$q50, total_predicted$q5, total_predicted$q95,
+    "Bayesian",    conf_int[2], conf_int[1], conf_int[3]
+)
+
+res[2,]$Q95 - res[2,]$Q5
+res[3,]$Q95 - res[3,]$Q5
+
+flextable(res)

### 2021-10-21 manuscript/figures.Rmd

@@
     keep_tex: yes
+header-includes: |
+  \pagenumbering{gobble}
 ---
@@
 
-```{r fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.", echo=FALSE, fig.width=1,out.height="4in"}
-knitr::include_graphics("../manuscript/figures/pareto_demo-1.pdf")
+```{r fig.cap="Censored lake area edf (solid line) and cdf estimate (dashed line).", echo=FALSE, fig.width=1,out.height="4in"}
+knitr::include_graphics("../manuscript/figures/predict_censor-1.pdf")
 ```
 
-```{r fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.", echo=FALSE, fig.width=1,out.height="4in"}
-knitr::include_graphics("../manuscript/figures/pareto_demo-1.pdf")
+```{r fig.cap="Stan model code", echo=FALSE, fig.width=1,out.height="5in"}
+knitr::include_graphics("../manuscript/figures/stan.pdf")
 ```
 
-```{r fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.", echo=FALSE, fig.width=1,out.height="4in"}
-knitr::include_graphics("../manuscript/figures/pareto_demo-1.pdf")
+```{r fig.cap="Median (black line) and central 95 percent interval estimates of alpha (red lines). Here the 'true' alpha is 0.9.", echo=FALSE, fig.width=1,out.height="4in"}
+knitr::include_graphics("../manuscript/figures/bayesian_model-1.pdf")
 ```
 
-```{r fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.", echo=FALSE, fig.width=1,out.height="4in"}
-knitr::include_graphics("../manuscript/figures/pareto_demo-1.pdf")
+```{r fig.cap="Median (black line) and central 95 percent interval estimates of total lake area (red lines).
+Here the true total lake area is marked with a dashed vertical line.", echo=FALSE, fig.width=1,out.height="4in"}
+knitr::include_graphics("../manuscript/figures/bayesian_area-1.pdf")
 ```

### 2021-10-28 manuscript/tables/reprex.R
formatted table 1

@@ -1,32 +1,0 @@
-library(reprex)
-
-tmp <- file.path(tempdir(), "foofy")
-dir.create(tmp)
-reprex({
-    #+ setup, include = FALSE
-    knitr::opts_chunk$set(results=FALSE)  
-"
-data {
-  int<lower=0> N;
-  real x[N];
-}
-parameters {
-  real<lower=0> alpha;
-  real<lower=0> theta;
-}
-model {
-  real lpa[N];
-
-  theta ~ gamma(1, 3);
-  alpha ~ gamma(1, 3);
-
-  for (i in 1:N) {
-    lpa[i] = pareto_lpdf(x[i] | theta, alpha);
-  }
-
-  target += sum(lpa);
-}
-"
-  
-  }, wd = ".", advertise = FALSE)
-list.files(dirname(tmp), pattern = "foofy")

### 2021-10-28 manuscript/tables/table_1.R

@@ -6,26 +6,37 @@ total_empirical <- sum(inv_cumulative_freq(cumulative_freq(y_raw)))
 
 res <- readRDS("data/res.rds")
 total_predicted <- data.frame(
-  q50 = sum(inv_cumulative_freq(res)), 
+  q50 = sum(inv_cumulative_freq(res)),
   q5 = sum(inv_cumulative_freq(
-    mutate(mutate(res, number_lower = coalesce(number_lower, number)), number = number_lower)
+    mutate(
+      mutate(res, number_lower = coalesce(number_lower, number)),
+      number = number_lower)
     )),
   q95 = sum(inv_cumulative_freq(
-    mutate(mutate(res, number_upper = coalesce(number_upper, number)), number = number_upper)
-    )) 
+    mutate(
+      mutate(res, number_upper = coalesce(number_upper, number)),
+      number = number_upper)
+    ))
   )
 
 area_bayes <- readRDS("data/area_bayes.rds")
 conf_int <- quantile(area_bayes, probs = c(0.025, 0.5, .975))
 
-res <- tibble::tribble(
-  ~ name,        ~ Q50, ~ Q5, ~ Q95,
-    "True",        total_empirical, NA, NA, 
-    "Frequentist", total_predicted$q50, total_predicted$q5, total_predicted$q95,
-    "Bayesian",    conf_int[2], conf_int[1], conf_int[3]
-)
+format_mil <- function(x){
+  round(x / 1000000, 3)
+}
+# format_mil(res[2,]$Q5)
 
-res[2,]$Q95 - res[2,]$Q5
-res[3,]$Q95 - res[3,]$Q5
+format_diff <- function(x){
+  round(x, 0)
+}
+# format_diff(conf_int[3] - conf_int[1])
 
-flextable(res)
+tibble::tribble(
+  ~ name,        ~ Q50, ~ Q5, ~ Q95, ~ Q95minusQ5,
+    "True",        format_mil(total_empirical), NA, NA, NA,
+    "Frequentist", format_mil(total_predicted$q50), format_mil(total_predicted$q5), format_mil(total_predicted$q95), format_diff(total_predicted$q95-total_predicted$q5),
+    "Bayesian",    format_mil(conf_int[2]), format_mil(conf_int[1]), format_mil(conf_int[3]), format_diff(conf_int[3]-conf_int[1])
+) %>%
+flextable() %>%
+colformat_num(na_str = "-")

### 2021-10-29 manuscript/figures.Rmd
stash figures pdf, closes #3

@@
 
-```{r fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.", echo=FALSE, fig.width=1,out.height="4in"}
+```{r fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function compared against equivalent calculations on the HydroLAKES dataset.", echo=FALSE, fig.width=1,out.height="4in"}
 knitr::include_graphics("../manuscript/figures/pareto_demo-1.pdf")
@@
 
-```{r fig.cap="Censored lake area edf (solid line) and cdf estimate (dashed line).", echo=FALSE, fig.width=1,out.height="4in"}
+```{r fig.cap="Censored lake area edf (solid line) and transformed cdf estimate (dashed line) compared against the HydroLAKES dataset.", echo=FALSE, fig.width=1,out.height="4in"}
 knitr::include_graphics("../manuscript/figures/predict_censor-1.pdf")
@@
 
-```{r fig.cap="Median (black line) and central 95 percent interval estimates of alpha (red lines). Here the 'true' alpha is 0.9.", echo=FALSE, fig.width=1,out.height="4in"}
+```{r fig.cap="Median (black line) and central 95 percent interval estimates of alpha (red lines).
+Here the 'true' alpha is 0.9 and is marked with a dashed line.", echo=FALSE, fig.width=1,out.height="4in"}
 knitr::include_graphics("../manuscript/figures/bayesian_model-1.pdf")

### 2022-09-26 scripts/03_bayesian.R
re-run with 0.1 km2 threshold

@@ -67,15 +67,22 @@ if (!file.exists("data/alphas.rds")) {
 alphas <- readRDS("data/alphas.rds")
 
 conf_int <- quantile(alphas, probs = c(0.025, 0.5, .975))
+alphas_sim <- rlnorm(n = 100000,
+  meanlog = log(conf_int[2]),
+  sdlog = conf_int[2] - conf_int[1])
+conf_int <- quantile(alphas_sim, probs = c(0.025, 0.5, .975))
+
 bayesian_model <- ggplot() +
-  geom_histogram(data = data.frame(alpha = alphas), aes(x = alpha)) +
+  geom_histogram(data = data.frame(alpha = alphas_sim),
+    aes(x = alpha), binwidth = (conf_int[2] - conf_int[1]) / 10) +
   geom_vline(aes(xintercept = conf_int[c(1, 3)]), color = "red") +
   geom_vline(aes(xintercept = conf_int[2])) +
   geom_vline(aes(xintercept = 0.9), linetype = 2) +
   xlim(
-    conf_int[2] - ((conf_int[2] - conf_int[1]) * 3),
-    conf_int[2] + ((conf_int[3] - conf_int[2]) * 3)
+    conf_int[2] - ((conf_int[2] - conf_int[1]) * 2.5),
+    conf_int[2] + ((conf_int[3] - conf_int[2]) * 2.5)
   )
+bayesian_model
 ggsave("manuscript/figures/bayesian_model-1.pdf", bayesian_model,
   width = 4.43, height = 2.33
 )
@@ -92,11 +99,11 @@ total_empirical <- sum(inv_cumulative_freq(cumulative_freq(y_raw)))
 if (!file.exists("data/area_bayes.rds")) {
   # find estimated density of censored lakes given alpha
   # back-out an estimate of total area
-  area_bayes <- sapply(alphas, function(a) {
+  area_bayes <- sapply(sample(alphas_sim, 700), function(a) {
     # a <- alphas[1]
     cf_extra_bayes <- select(cf_extra, area)
     cf_extra_bayes$density <- (log(cf_extra_bayes$area) * (a * -1)) +
-      log(min(y_censored)) - 0.1
+      log(min(y_censored)) - 0.085
     cf_extra_bayes$type <- "predicted_bayes"
     cf_extra_bayes$number <- exp(cf_extra_bayes$density + max(log(cf$number)))
[…]

### 2022-09-27 scripts/03_bayesian.R
fight with convergence

@@ -50,6 +50,7 @@ if (!file.exists("data/pareto_bayes.rds")) {
     file = "pareto_model.stan",
     data = list(N = length(y_censored), x = y_censored),
     iter = 25000,
+    chains=4,
     control = list(adapt_delta = 0.9),
     init = init_f
   )
@@ -96,7 +97,7 @@ if (!file.exists("data/area_bayes.rds")) {
     # a <- alphas[1]
     cf_extra_bayes <- select(cf_extra, area)
     cf_extra_bayes$density <- (log(cf_extra_bayes$area) * (a * -1)) +
-      log(min(y_censored)) - 0.1
+      log(min(y_censored)) - 0.085
     cf_extra_bayes$type <- "predicted_bayes"
     cf_extra_bayes$number <- exp(cf_extra_bayes$density + max(log(cf$number)))
     res <- dplyr::bind_rows(cf_extra_bayes, cf)

### 2022-12-14 manuscript/figures.Rmd
more specific description of dashed line

@@
 
-```{r fig.cap="Censored lake area edf (solid line) and transformed cdf estimate (dashed line) compared against the HydroLAKES dataset.", echo=FALSE, fig.width=1,out.height="4in"}
+```{r fig.cap="Simulated lake area (blue line) where the portion (>= 0.1 $km^2$) calculated from the edf is denoted by a solid line and the portion (< 0.1 $km^2$) calculated from the transformed cdf is denoted by a dashed line.
+For comparison, the edf of the HydroLAKES dataset (red line) is also shown.", echo=FALSE, fig.width=1,out.height="4in"}
 knitr::include_graphics("../manuscript/figures/predict_censor-1.pdf")
-```
-
-```{r fig.cap="Stan model code", echo=FALSE, fig.width=1,out.height="5in"}
-knitr::include_graphics("../manuscript/figures/stan.pdf")
 ```

### 2022-12-19 scripts/00_simulation.R
make figs 1+2 colorblind friendly

@@ -61,7 +61,8 @@ gg_individual <- rbind(y$y_individual, hydrolakes$hydrolakes_individual) %>%
   ylab("log(n)") +
   xlab("log(area)") +
   theme_minimal() +
-  theme(legend.position = "none")
+  theme(legend.position = "none") +
+  scale_color_viridis_d(direction=-1)
 
 gg_cumulative <- rbind(y$y_cumulative, hydrolakes$hydrolakes_cumulative) %>%
   ggplot() +
@@ -71,7 +72,8 @@ gg_cumulative <- rbind(y$y_cumulative, hydrolakes$hydrolakes_cumulative) %>%
   scale_y_log10() +
   ylab("n > area") +
   xlab("area") +
-  labs(color = "")
+  labs(color = "") +
+  scale_color_viridis_d(direction=-1)
 
 pareto_demo <- plot_grid(gg_individual, gg_cumulative,
   nrow = 1, rel_widths = c(0.68, 1)

### 2022-12-19 scripts/01_censoring.R

@@ -95,7 +95,8 @@ predict_censor <- rbind(
   geom_line(aes(area, number, color = name, linetype = type)) +
   theme_minimal() + scale_x_log10() + scale_y_log10() +
   ylab("n > area") + xlab("area") +
-  labs(color = "")
+  labs(color = "") +
+  scale_color_viridis_d(direction=-1)
 
 ggsave("manuscript/figures/predict_censor-1.pdf", predict_censor,
   width = 4.28, height = 2.33)

### 2022-12-19 manuscript/figures.Rmd
more dashed line explanation

--- 2021-10-29 manuscript/figures.Rmd
@@
 
-```{r fig.cap="Censored lake area edf (solid line) and transformed cdf estimate (dashed line) compared against the HydroLAKES dataset.", echo=FALSE, fig.width=1,out.height="4in"}
+```{r fig.cap="Simulated lake area (blue line) where the portion calculated from the edf (>= 0.1 $km^2$) is denoted by a solid line and the portion calculated from the transformed cdf (< 0.1 $km^2$) is denoted by a dashed line.
+For comparison, the edf of the HydroLAKES dataset (solid red line) is also shown.", echo=FALSE, fig.width=1,out.height="4in"}
 knitr::include_graphics("../manuscript/figures/predict_censor-1.pdf")
-```
-
-```{r fig.cap="Stan model code", echo=FALSE, fig.width=1,out.height="5in"}
-knitr::include_graphics("../manuscript/figures/stan.pdf")
 ```

### 2022-12-19 manuscript/manuscript.tex
towards syncing with word file
syncing pdf with docx

@@
 \usepackage{amsfonts}       % blackboard math symbols
-%\usepackage{nicefrac}      % compact symbols for 1/2, etc.
 \usepackage{microtype}      % microtypography
@@
 
-\author{ \href{https://orcid.org/0000-0002-5924-2464}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}J.~Stachelek} \\
+\author{ \href{https://orcid.org/0000-0002-5924-2464}{\includegraphics[scale=0.06]{orcid.pdf}\hspace{1mm}Jemma Stachelek} \\
 	Earth System Observations\\
@@
 pdfsubject={stat.AP},
-pdfauthor={J.~Stachelek}
+pdfauthor={Jemma Stachelek}
 }
@@
 \begin{abstract}
-	Size is a critical factor determining the rate and occurrence of specific lake processes such as carbon sequestration and greenhouse gas emissions.
-Because we do not have a complete census of all lakes, upscaling estimates of such processes at broad spatial scales requires the use of lake size-abundance distributions rather than empirical measurements of area.
-Existing lake census efforts are incomplete such that as lakes become smaller they are more likely to be omitted either because they are too small to be resolved from remote sensing products or because of limited ground surveying effort (i.e.
+	Size is a critical factor determining the rate and occurrence of specific lake processes such as carbon sequestration and greenhouse gas emissions and emerging evidence suggests that small lakes in particular have particularly large $CO^2$ flux rates.
+Because we do not have a complete census of all lakes, upscaling estimates of such processes to small lakes at broad spatial scales requires the use of lake size-abundance distributions rather than empirical measurements of area.
+Existing lake census efforts are incomplete such that as lakes become smaller, they are more likely to be omitted either because they are too small to be resolved from remote sensing products or because of limited ground surveying effort (i.e.
 "censoring" of small lakes relative to large lakes).
 The present study explores one potential shortcoming of prior approaches estimating global lake area using lake size-abundance distributions.
-Namely, these prior approaches rely on frequentist curve fitting techniques combined with an ad-hoc cutoff determination strategy (visual inspection to determine a likely censoring point).
-This yields an over-exact lake area estimate that is overwhelmingly reported with no uncertainty bounds.
-I address this shortcoming by fitting models in a Bayesian framework where each parameter contributes uncertainty to model estimates.
-I show that such although such models produce a more realistic estimate of lake area uncertainty they underestimate true total lake area.
-The degree of this underestimation is likely related to the proportion of the dataset subject to censoring.
-Ultimately, this may explain the fact that total lake area estimates have increased through time as the resolution of lake databases has improved.
+Namely, that these prior approaches rely on frequentist curve fitting techniques combined with an ad-hoc cutoff determination strategy (visual inspection to determine a likely censoring point).
+This yields an over-exact lake area estimate that is typically reported with no uncertainty bounds.
+I show how these shortcomings can be addressed with a Bayesian model that produces larger estimates of lake area uncertainty relative to the typical approach.
+When used as part of a sensitivity analysis, such an approach has the potential to enable more robust intercomparisons among studies of aquatic processes upscaling.
 \end{abstract}
@@
 \section{Introduction}
-Area is a critical factor determining the contribution of lakes to global biogeochemical cycling.
-As such, existing evidence suggests that lakes are key components of overall carbon burial, sequestration, and greenhouse gas emissions \citep{delsontroGreenhouseGasEmissions2018, kellerGlobalCarbonBudget2021}.
+Size is a critical factor determining the contribution of lakes to global biogeochemical cycling.
+As such, existing evidence suggests that lakes are key components of overall carbon burial, sequestration, and greenhouse gas emissions with small lakes having particularly high $CO^2$ flux rates \citep{delsontroGreenhouseGasEmissions2018, kellerGlobalCarbonBudget2021}.
+When upscaled to broad spatial extents, the high flux rates of these small lakes translate to a disproportionately large influence on overall $CO^2$ emissions \citep{piMappingGlobalLake2022}.
 The calculation of such fluxes requires two terms 1) an areal flux rate and 2) the distribution of total lake area.
-As with all estimates at broad spatial scales, uncertainty estimates are of great interest both because they allow for intercomparisons among similar lake studies and because they allow for comparisons between lakes and other ecosystems.
+As with all quantitative estimates at broad spatial scales, uncertainty estimates are of great interest both because they allow for intercomparisons among similar studies and because they allow for comparisons between ecosystems.
 
-The techniques typically used to determine uncertainty in areal flux rates and the techniques typically used to determine uncertainty in total lake area distributions differ greatly.
+The techniques typically used to determine uncertainty in areal flux rates differ greatly from those used to determine uncertainty in total lake area distributions.
 In the former case, uncertainties can typically be calculated in straightforward manner from literature compilations \citep{delsontroGreenhouseGasEmissions2018, kellerGlobalCarbonBudget2021}.
 In the latter case, uncertainty in total lake area distributions is challenged by the fact that no existing database is a complete census of all lakes \citep{messagerEstimatingVolumeAge2016} and the distribution of individual lake areas spans a range of approximately 7 orders of magnitude.
-The largest lakes are so big they could otherwise be classified as inland seas (> $10^4$ $km^2$) while the smallest are barely larger than a regulation sized soccer (football) field (> $10^{-3}$ $km^2$).
+The largest lakes are so big they could otherwise be classified as inland seas (> 104 $km^2$) while the smallest are barely larger than a regulation sized soccer (football) field (> $10^{-3}$ $km^2$).
+
 A consequence of these challenges is that uncertainty varies within the distribution as a function of lake size.
-The area of the largest lakes is known exactly while the area of the smallest lakes is incomplete below a certain unknown threshold.
-This can occur either because lakes are too small to be resolved from remote sensing products or because of limited ground surveying effort.
-The omission or "censoring" of small lakes occurs because we know that small lakes exist but below a certain threshold we have limited knowledge of their exact areas \citep{hamiltonEstimationFractalDimension1992}.
+The area of the largest lakes is known with a high degree of certainty while the area of the smallest lakes is unknown below a certain unknown threshold.
+The area of a small lake can be unknown either because it is too small to be resolved from remote sensing products or because of limited ground surveying effort.
+The omission or "censoring" of small lakes occurs because we know that small lakes exist but below a certain threshold, we have limited knowledge of their exact areas \citep{hamiltonEstimationFractalDimension1992}.
+Such censoring errors may have an outsized impact on upscaled estimates of aquatic processes such as $CO^2$ flux because small lakes have been shown to have particularly large $CO^2$ flux rates \citep{delsontroGreenhouseGasEmissions2018,piMappingGlobalLake2022}.
 
@@
 As large lakes become even larger, they have a greater probability of intersecting a continent edge and becoming an estuary or embayment rather than a lake.
-Consquently, lake databases are said to be truncated on large lakes because we know that large lakes are essentially fixed in space and cannot occur in any given location \citep{hamiltonEstimationFractalDimension1992}.
+Consequently, lake databases are said to be truncated on large lakes because we know that large lakes are essentially fixed in space and cannot occur in any given location \citep{hamiltonEstimationFractalDimension1992}.
 
-From the preceding discussion it is clear that estimating total lake area requires a method of dealing both with the fact that lake databases are 1) truncated at large lakes and 2) censored at small lakes.
-Previous studies estimating global lake area have essentially not dealt with the first issue but instead modified their estimation process to limit its effect on the results (see methods section).
-They have dealt with the second issue by specifying an ad-hoc cutoff value below which empirically determined lake areas are discarded and subsequently back calculated.
+Given that no existing database is a complete census of all lakes, yet we have near exact estimates of the area of large lakes, estimating total lake area requires a method of dealing both with the fact that lake databases are 1) truncated at large lakes and 2) censored at small lakes.
+Prior studies estimating global lake area have mostly not dealt with the first issue (but see \citet{seekell2011does}).
+Note that in some instances the estimation process can be modified to limit its effect on the results (see Methods section).
+In contrast to the first issue, prior studies have typically addressed the second issue by specifying an ad-hoc cutoff value below which empirically determined lake areas are discarded and subsequently back-calculated.
 Hereafter, I refer to this strategy as the "cutoff" method.
@@
 The use of a simulated dataset has the further benefit of avoiding potential confounding factors such as heterogeneity of survey effort or unknown data precision.
+The following analysis assumes that total global lake area has minimal temporal variability or at least minimal spatial trend.
+As a result, reported total lake areas can be thought of as “anticipated” or equilibrium long-run lake area rather than true totals representing any specific point in time.
 
 \section{Methods}
+
+\subsection{Data description}
+
+I compared the properties of a simulated dataset of lake areas (described below) against that of the HydroLAKES dataset \citep{messagerEstimatingVolumeAge2016}.
+HydroLAKES was created as a compilation of existing broad scale lake datasets including the SRTM Water Body Data \citep{slater2006srtm} and contains information on lakes between 0.1 and 500 $km^2$.
+The total area of the HydroLAKES dataset is 2.67 mil.
+$km^2$.
+
+\subsection{Model overview}
 
@@
 
-For evaluation purposes, I generated a simulated dataset of 10,000 lake areas following the Pareto distribution using inverse transform sampling \citep{newman_power_2005}.
-Lakes in my simulated dataset have a minimum and maximum area of approximately 1 and 81000 $km^2$ respectively.
-This maximum was chosen to be approximately as large as Lake Superior but less than the Caspian Sea following \citet{lehnerDevelopmentValidationGlobal2004}.
-The "true" total area of these lakes is approximately 230000 $km^2$.
-I simulated a censored lake dataset by excluding lakes smaller than $e^{1}$.
-This excludes (i.e.
-censors) approximately 60\% of the total dataset.
+For evaluation purposes, I generated a simulated dataset of 153,000 lake areas following the Pareto distribution using inverse transform sampling \citep{newman_power_2005}.
+Lakes in my simulated dataset have a minimum and maximum area of approximately 0.1 and 81,000 $km^2$ respectively.
+This maximum was chosen to be approximately as large as Lake Superior but less than the Caspian Sea following \citep{lehnerDevelopmentValidationGlobal2004}.
+The number of lakes in the simulated dataset was adjusted so that the total “true” area of the dataset (3.28 mil $km^2$) would approximately match (but still exceed) the total reported in the HydroLAKES dataset (2.93 mil $km^2$).
+I simulated a censored lake dataset by excluding approximately 60\% of lakes in the total dataset.
 I approximated the "true" lake area total by constructing the empirical distribution function (edf) of the data which approximates the underlying Pareto cdf \citep{newman_power_2005}.
 Then I used this estimate of the cdf slope to generate cdf estimates for the censored lakes.
-I combined these cdf estimates with the edf values from the known lakes before calculating the sum of the inverted distribution (Figure \ref{fig:predict_area}).
+I combined these cdf estimates with the edf values from the known lakes before calculating the sum of the inverted distribution \ref{fig:predict_area}.
 

### 2022-12-19 manuscript/manuscript.tex
sync docx and pdf

@@
 
-\begin{figure}
+\begin{figure}[h]
 	\centering
 	\includegraphics{figures/pareto_demo-1}
-	\caption{Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.}
+	\caption{Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function compared against equivalent calculations on the HydroLAKES dataset.}
 	\label{fig:pareto_demo}
@@
 I evaluated uncertainty in both $a$ and total lake area in a Bayesian framework using Stan \citep{standevelopmentteamStanModelingLanguage2017}.
-Instead of computing on the edf (as in the frequentist case), I computed directly on the pdf with the following Stan model:
+Instead of computing on the edf (as in the frequentist case), I computed directly on the pdf with a Stan model \citep
+[See][]{jemma_stachelek_2022_7459227}.
+I used uninformative gamma priors for both the $x_{min}$ and $a$ parameters following \citet{scollnikCompositeLognormalParetoModels2007}.
+I ran the model with four chains of 25,000 iterations and used the Stan defaults for thinning and burn-in which specify no thinning and discarding the first half (12,500) of iterations.
+I examined model fits to ensure that all models had acceptable convergence of MCMC chains.
 
-\begin{minipage}[c]{0.95\textwidth}
-	\begin{lstlisting}
-data {
-  int<lower=0> N;
-  real x[N];
-}
-parameters {
-  real<lower=0> alpha;
-  real<lower=0> xmin;
-}
-model {
-  real lpa[N];
+Finally, I evaluated empirical uncertainty in total lake area using the size-dependent detection/non-detection estimates reported by \citet{cheruvelil2021lagos} which found that about 80\% of lakes $<$ 1 $km^2$ and 95\% of lakes between 1 and 10 $km^2$ were accurately represented in the US National Hydrograph Network.
+I implemented a random sampling (n=500) procedure using these detection fractions whereby empirical uncertainty is reported as the quantiles of the resulting distribution.
 
-  xmin ~ gamma(.001, .001);
-  alpha ~ gamma(.001, .001);
+\section{Assessment}
 
-  for (i in 1:N) {
-    lpa[i] = pareto_lpdf(x[i] | xmin, alpha);
-  }
+The total area of the simulated dataset (3.28 mil.
+$km^2$, Table \ref{tab:area}) was larger than the empirical total of the HydroLAKES (2.67 mil.
+$km^2$) dataset.
+Visual inspection of the frequentist results, which are computed on the edf, appeared to produce a reasonable density estimate for small, censored lakes (Figure \ref{fig:predict_censor}).
+Although frequentist estimates of total lake area were close to the "true" value, uncertainty bounds were unrealistically small compared to an empirical estimate of uncertainty (Table \ref{tab:area}).
 
-  target += sum(lpa);
-}
-\end{lstlisting}
-\end{minipage}
-
-I used uninformative gamma priors for both the $x_{min}$ and $a$ parameters following \citet{scollnikCompositeLognormalParetoModels2007}.
-I ran the model with four chains of 8,000 iterations and used the Stan defaults for burn-in and thinning which specify a burn-in of half the iterations and no thinning.
-
-\section{Results}
-
-Visual inspection of the frequentist method of computing on the edf appeared to produce a reasonable density estimate for small censored lakes (Figure \ref{fig:predict_area}).
-In addition, estimates of total lake area are somewhat close to the "true" value (Table \ref{tab:pred_table}).
-However, uncertainty around the frequentist estimates is unrealistically small (Figure \ref{tab:frequentist_uncertainty}).
+In contrast to the fixed a and fixed total lake area estimates produced by the frequentist approach, I found substantial variability in both a (95\% CI: 0.65, 1.02) and total area using a Bayesian approach (Figure \ref{fig:bayesian_model}, \ref{fig:bayesian_area}).
+In particular, the Bayesian 95\% credible intervals for a encapsulated the true value (Figure \ref{fig:bayesian_model}).
+Despite larger uncertainty estimates using a Bayesian approach (i.e.
+more closely matching the empirical estimate of uncertainty), the true value of total lake area was underestimated.
+The reason for this underestimation is likely owed to incomplete “back calculation” in the cutoff method (i.e.
+back calculated lakes did not extend all the way to the smallest lakes in the simulated dataset).
 
@@
 	\includegraphics{figures/predict_censor-1}
-	\caption{Censored lake area edf (solid line) and cdf estimate (dashed line).}
+	\caption{Simulated lake area (blue line) where the portion calculated from the edf (>= 0.1 $km^2$) is denoted by a solid line and the portion calculated from the transformed cdf (< 0.1 $km^2$) is denoted by a dashed line.
+For comparison, the edf of the HydroLAKES dataset (solid red line) is also shown.}
 	\label{fig:predict_censor}
@@
 
-% \begin{figure}
-% 	\centering
-% 	\includegraphics{figures/frequentist_uncertainty-1}
-% 	\caption{Confidence interval of frequentist density predictions for censored data (the dashed portion in Figure 2. The confidence interval here has essentially no width.}
-% 	\label{tab:frequentist_uncertainty}
-% \end{figure}
 \begin{table}
-	\caption{Frequentist uncertainty}
+	\caption{Comparison between the true total lake area in the simulated dataset against frequentist and Bayesian estimates.
+The true total is a point estimate while uncertainty is displayed via differences between the 95th and 5th quantile.}
 	\centering
-	\begin{tabular}{ccc}
+	\begin{tabular}{lllll}
 		\toprule
-		Q5 ($km^2$) & Q50 ($km^2$) & Q95 ($km^2$) \\
+		& Q50 (mil $km^2$) & Q5 (mil $km^2$) & Q95 (mil $km^2$) & Q95-Q5 \\
 		\midrule
-		2.271e+05          & 2.263e+05   & 123     \\
+		True (point estimate) & 3.278 & - & - & - \\
+		True (80\% non-detect) & 3.205 & 3.205 & 3.205 & 673  \\
+		Frequentist & 3.274 & 3.274 & 3.274 & 14  \\
+		Bayesian & 3.199 & 3.158 & 3.273 & 115172 \\
 		\bottomrule
 	   \end{tabular}	
-	\label{tab:frequentist_uncertainty}
+	\label{tab:area}
 \end{table}
@@
 
-Instead of the essentially fixed $a$ and total lake area estimates produced by the frequentist approach, I found substantial variability in both $a$ (95\% CI: 0.86, 0.92) and total area using a Bayesian approach (Fig \ref{fig:bayesian_model}, \ref{fig:bayesian_area}).
-In particular, the Bayesian 95\% credible intervals for both $a$ and total lake area encapsulate the true values (Fig \ref{fig:bayesian_model}, \ref{fig:bayesian_area}).
-Despite better uncertainty estimates using a Bayesian approach, both the frequentist and Bayesian approaches underestimated the true value of $a$ and total lake area.
-
-\begin{table}
-	\caption{'True' total lake area from the uncensored edf and estimated lake area from a combination of the censored edf and the estimated cdf.}
-	\centering
-	\begin{tabular}{cc}
-		\toprule
-		True Area ($km^2$) & Estimated Area ($km^2$) \\
-		\midrule
-		2.271e+05          & 2.263e+05               \\
-		\bottomrule
-	\end{tabular}
-	\label{tab:pred_table}
-\end{table}
-
 \section{Discussion}
 
-I have shown that the typical frequentist cutoff method produces reasonable estimates of the density of small censored lakes but that it does not capture uncertainty in total lake area (Table \ref{tab:pred_table}, Figure \ref{tab:frequentist_uncertainty}).
-Furthermore, I have shown that models fit using a Bayesian approach indicate substantial uncertainty in both total lake area and the underlying Pareto shape parameter $a$ used to derive these estimates (Figure \ref{fig:bayesian_model}, \ref{fig:bayesian_area}).
+I have shown that the even with realistic estimates of the density of small censored lakes, the typical frequentist cutoff method does not reasonably capture uncertainty in total lake area (Table \ref{tab:area}).
+However, the results of the Bayesian approach by contrast, indicate substantial uncertainty in both total lake area and the underlying Pareto shape parameter a used to derive these estimates (Figure \ref{fig:bayesian_model}, \ref{fig:bayesian_area}).
+The combination of accurate lake area accounting and realistic uncertainty estimates offered by this approach may provide a more principled way to assess the relative influence of different lake area classes on global lake fluxes ($CO^2$, $CH^4$, etc.)  beyond ad-hoc study intercomparisons or even meta-analysis on collections of prior studies.
 
-Although the 95\% credible interval of the Bayesian total lake area estimates encapsulate the true total lake area, the median value underestimates the true total lake area (Figure \ref{fig:bayesian_area}).
-It is likely that such underestimation, would increase with a greater proportion of censoring.
-This may explain the steady increase in estimates of global lake area through time from approximately 3 to 5 million $km^{2}$ as lake area databases have improved their accuracy \citep{lehnerDevelopmentValidationGlobal2004, downingGlobalAbundanceSize2006, verpoorterGlobalInventoryLakes2014}.
-Future work on estimating global lake area should consider implementing a sensitivity analysis looking at the response of total area estimates to variation in the degree of censoring.
+The Bayesian model estimate of total lake area was an underestimate of the true total lake area of the simulated datasets (Figure \ref{fig:bayesian_area}).
+Some of this underestimation is because the Pareto shape parameter a was estimated from the truncated dataset (to simulate the fact that observed datasets such as HydroLAKES have some unknown degree of truncation) and was lower ({\raise.17ex\hbox{$\scriptstyle\mathtt{\sim}$}}0.82) than the true value of 0.9.
+This lower value of a resulted in lower than expected total lake area.
+Such underestimation would increase with a greater proportion of censoring.
+Another likely contributor to lake area underestimation is differences in the properties of the simulated data compared to an empirical dataset like HydroLAKES.
+Note that the simulated dataset diverges notably from HydroLAKES for the largest waterbodies because these are few in number such that randomness in the simulation is more obvious (Figure \ref{fig:pareto_demo}).
+Taken together, these factors (i.e.
+less and less truncation and better data on the largest waterbodies) may explain the steady increase in estimates of global lake area through time from approximately 3 to 5 million $km^2$ as lake area databases have improved their accuracy \citep{downingGlobalAbundanceSize2006,lehnerDevelopmentValidationGlobal2004,verpoorterGlobalInventoryLakes2014}.
+Future work on estimating global lake area may consider implementing a sensitivity analysis looking at the response of total area estimates to variation in the degree of censoring.
+
+Another area where further sensitivity analyses may be warranted is in the generation of total lake area estimates from Pareto realizations.
+Typically this is done a single time such that any particular reported estimate comes only from a single realization.
+To generate the simulated dataset for the present study, I generated multiple realizations with the goal of arriving at one that approximately matched the total reported in HydroLAKES.
+Although each realization matched the Pareto properties of HydroLAKES, the sum of many of the realizations differed markedly from each other and from HydroLAKES.
+This demonstrates the need for generating multiple Pareto estimates of total lake area for aquatic process upscaling as part of a sensitivity analysis rather than taking only a single realization.
 
 A caveat of the present study is that it is difficult to confirm whether or not any particular dataset follows a fractal generating or Pareto process without strong prior knowledge.
-Indeed data following many heavy-tailed distributions such as the lognormal or negative exponential can appear to be equivalent to power-law distributions such as the Pareto \citep{clausetPowerlawDistributionsEmpirical2009}.
+As a result, there is some degree to which we cannot know whether the distribution of lake areas truly aligns with the Pareto distribution \citep{seekell2011does}.
+Indeed data following many heavy-tailed distributions such as the lognormal or negative exponential can appear to be equivalent to power-law distributions such as the Pareto \citep{clausetPowerlawDistributionsEmpirical2009,seekell2011does}.
 The results of the present study confirm that an apparent change in the shape of the lower tail of a lake area distribution does not necessarily indicate a change in the data generating process.
-Note that the simulation dataset analyzed herein \emph{is} a truly fractal generated dataset with a homogeneous data generating process yet a cursory look would seem to indicate a change of data-generating process in the lower tail (Figure \ref{fig:predict_area}).
+Note that the simulation dataset analyzed herein is a truly fractal generated dataset with a homogeneous data generating process, yet a cursory look would seem to indicate a change of data-generating process in the lower tail (Figure \ref{fig:predict_censor}).
 
 In addition to a sensitivity analysis of censoring, future work might consider more complex models that treat lake areas as a mixture of a Pareto distribution for small lakes and either a negative exponential or lognormal distribution for large lakes.
-Such an approach has been demonstrated by \citet{bonabeau_scaling_1999-1} and \citet{scollnikCompositeLognormalParetoModels2007}.
+Such an approach has been demonstrated by \citep{bonabeau_scaling_1999-1} and \citep{scollnikCompositeLognormalParetoModels2007}.
 Both studies show that the point at which the distribution mixtures converge can provide valuable inference.
-In the case of lakes, such a convergence point may indicate a change in the data generating process such as the point at which lake areas are controlled by continent placement rather than fractal landscape morphology \citep{goodchildLakesFractalSurfaces1988, hamiltonEstimationFractalDimension1992}.
+In the case of lakes, such a convergence point may indicate a change in the data generating process such as the point at which lake areas are controlled by continent placement rather than fractal landscape morphology \citep{goodchildLakesFractalSurfaces1988,hamiltonEstimationFractalDimension1992}.
+Knowing such a convergence point would provide a data-driven estimate of what constitutes a “large lake” beyond ad-hoc cutoffs and supplement existing definitions based on hydrodynamics and circulation.
+
+% \begin{minipage}[c]{0.95\textwidth}
+% 	\begin{lstlisting}
+% data {
+%   int<lower=0> N;
+%   real x[N];
+% }
+% parameters {
+%   real<lower=0> alpha;
+%   real<lower=0> xmin;
+% }
+% model {
+%   real lpa[N];
+
+%   xmin ~ gamma(.001, .001);
+%   alpha ~ gamma(.001, .001);
+
+%   for (i in 1:N) {
+%     lpa[i] = pareto_lpdf(x[i] | xmin, alpha);
+%   }
+
+%   target += sum(lpa);
+% }
+% \end{lstlisting}
+% \end{minipage}
 

### 2022-12-19 manuscript/figures.Rmd
fix color typo in fig caption

--- 2021-10-29 manuscript/figures.Rmd
@@
 
-```{r fig.cap="Censored lake area edf (solid line) and transformed cdf estimate (dashed line) compared against the HydroLAKES dataset.", echo=FALSE, fig.width=1,out.height="4in"}
+```{r fig.cap="Simulated lake area (purple line) where the portion calculated from the edf (>= 0.1 $km^2$) is denoted by a solid line and the portion calculated from the transformed cdf (< 0.1 $km^2$) is denoted by a dashed line.
+For comparison, the edf of the HydroLAKES dataset (solid yellow line) is also shown.", echo=FALSE, fig.width=1,out.height="4in"}
 knitr::include_graphics("../manuscript/figures/predict_censor-1.pdf")
-```
-
-```{r fig.cap="Stan model code", echo=FALSE, fig.width=1,out.height="5in"}
-knitr::include_graphics("../manuscript/figures/stan.pdf")
 ```

### 2022-12-19 manuscript/manuscript.tex
fix color typo in fig caption

@@
 	\includegraphics{figures/predict_censor-1}
-	\caption{Simulated lake area (blue line) where the portion calculated from the edf (>= 0.1 $km^2$) is denoted by a solid line and the portion calculated from the transformed cdf (< 0.1 $km^2$) is denoted by a dashed line.
-For comparison, the edf of the HydroLAKES dataset (solid red line) is also shown.}
+	\caption{Simulated lake area (purple line) where the portion calculated from the edf (>= 0.1 $km^2$) is denoted by a solid line and the portion calculated from the transformed cdf (< 0.1 $km^2$) is denoted by a dashed line.
+For comparison, the edf of the HydroLAKES dataset (solid yellow line) is also shown.}
 	\label{fig:predict_censor}

### 2023-01-30 manuscript/manuscript.tex
prep for accepted release

@@
 % Uncomment to override  the `A preprint' in the header
-%\renewcommand{\headeright}{Technical Report}
-%\renewcommand{\undertitle}{Technical Report}
+\renewcommand{\headeright}{\href{https://doi.org/}{DOI: 10.XXX/XXX}}
+\renewcommand{\undertitle}{This is an Accepted Manuscript of an article published in Limnology \& Oceanography: Methods \href{https://doi.org/}{DOI: 10.XXX/XXX}}
 \renewcommand{\shorttitle}{Pareto lake area uncertainty}
@@
 
+\section{Data Availability Statement}
+
+All data and code associated with this manuscript are available at: \href{https://doi.org/10.5281/zenodo.7459226}{https://doi.org/10.5281/zenodo.7459226}
+
+\section{Acknowledgements}
+
+This work was supported by the U.S.
+National Science Foundation’s Harnessing the Data Revolution Program (OAC-1934633) and Los Alamos National Laboratory (LDRD-20210213ER, LDRD-20220697PRD1).
+I thank the following individuals for a friendly review of an earlier draft: P.C.
+Hanson, E.
+Zipkin.
+
 % \begin{minipage}[c]{0.95\textwidth}

### 2023-02-06 manuscript/manuscript.tex
add paper doi

@@
 % Uncomment to override  the `A preprint' in the header
-\renewcommand{\headeright}{\href{https://doi.org/}{DOI: 10.XXX/XXX}}
-\renewcommand{\undertitle}{This is an Accepted Manuscript of an article published in Limnology \& Oceanography: Methods \href{https://doi.org/}{DOI: 10.XXX/XXX}}
+\renewcommand{\headeright}{\href{https://doi.org/}{DOI: 10.1002/LOM3.10536}}
+\renewcommand{\undertitle}{This is an Accepted Manuscript of an article published in Limnology \& Oceanography: Methods \href{https://doi.org/10.1002/LOM3.10536}{DOI: 10.1002/LOM3.10536}}
 \renewcommand{\shorttitle}{Pareto lake area uncertainty}

### 2023-02-06 manuscript/manuscript.tex
typo fix

@@
 	\includegraphics{figures/predict_censor-1}
-	\caption{Simulated lake area (purple line) where the portion calculated from the edf (>= 0.1 $km^2$) is denoted by a solid line and the portion calculated from the transformed cdf (< 0.1 $km^2$) is denoted by a dashed line.
-For comparison, the edf of the HydroLAKES dataset (solid yellow line) is also shown.}
+	\caption{Simulated lake area (yellow line) where the portion calculated from the edf (>= 0.1 $km^2$) is denoted by a solid line and the portion calculated from the transformed cdf (< 0.1 $km^2$) is denoted by a dashed line.
+For comparison, the edf of the HydroLAKES dataset (solid purple line) is also shown.}
 	\label{fig:predict_censor}

### 2023-02-06 manuscript/manuscript.tex
figure link fix

@@
 Then I used this estimate of the cdf slope to generate cdf estimates for the censored lakes.
-I combined these cdf estimates with the edf values from the known lakes before calculating the sum of the inverted distribution \ref{fig:predict_area}.
+I combined these cdf estimates with the edf values from the known lakes before calculating the sum of the inverted distribution \ref{fig:predict_censor}.
 
